### 2011-10-04 old/CacheProof.tex
retaining old stuff

\documentclass{article}

\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\begin{document}

\begin{lemma}
Child-state is at x due to parent receiving x-resp and child is at y due to child receiving y-resp $\Rightarrow$ x = y.
\label{CRecvPRecvNetwork}
\end{lemma}

\begin{proof}
Proof by contradiction, assuming x $\neq$ y.

\begin{figure}
\centering
%\includegraphics{CRecvPRecvNetwork.1}
\end{figure}

Child has sent x-resp before receiving y-resp ($\because$ y-resp is the last message causing child transition).

Parent has sent y-resp before receiving x-resp ($\because$ x-resp is the last message causing child-state transition).

$\Rightarrow$ The corresponding y-req should have been sent by the child before sending x-resp ($\because$ FIFO ordering between child to parent req, resp).

$\Rightarrow$ x-resp is not voluntary.

$\Rightarrow$ x-req is sent by parent before sending y-resp ($\because$ x-resp is sent by child, and hence x-req is received by child before receving y-resp and FIFO order ordering between parent to child messages).

But y-resp will not be sent by parent if there is a pending request.

\begin{figure}
\centering
%\includegraphics{CRecvPRecvNetwork.2}
\end{figure}

$\Rightarrow \Leftarrow$
\end{proof}

\begin{lemma}
Child-state is at x due to parent sending x-resp and child is at y due to child sending y-resp, corresponding x-resp was sent by parent before receiving y-resp and y-resp was sent by child before receiving x-resp $\Rightarrow$ x = y.
\label{InsaneNetwork}
\end{lemma}

\begin{proof}
Proof by contradiction, assuming x $\neq$ y.

\begin{figure}
\centering
%\includegraphics{InsaneNetwork.1}
\end{figure}

x-resp has not reached child ($\because$ y-resp is the last message sent by child causing child transition, and child is still in y).

y-resp has not reached parent ($\because$ x-resp is the last message sent by parent causing child-state transition, and child-state is still in x).

Corresponding x-req was received by parent before receiving y-resp.

$\Rightarrow$ x-req was sent by child before sending y-resp ($\because$ FIFO ordering between child to parent req, resp).

$\Rightarrow$ y-resp is not voluntary.

$\Rightarrow$ Corresponding y-req was sent by parent before sending x-resp ($\because$ FIFO ordering between parent to child messages).

$\Rightarrow$ x-resp was sent by parent before receiving y-resp for y-req sent by parent which is not possible ($\because$ parent does not respond when there is a pending request).

\begin{figure}
\centering
%\includegraphics{InsaneNetwork.2}
\end{figure}

$\Rightarrow \Leftarrow$

\end{proof}

\begin{lemma}
Child-state is at x due to parent sending x-resp and child is at y due to child receiving y-resp $\Rightarrow$ y $\le$ x.
\label{CRecvPSend<=}
\end{lemma}

\begin{proof}
Proof by contradiction, assuming x $<$ y.

\begin{figure}
\centering
%\includegraphics{CRecvPSend<=.1}
\end{figure}

Parent sends x-resp only after y-resp ($\because$ x-resp is the last message causing child-state transition).

Child can send corresponding x-req for the x-resp from parent only after child receives y-resp ($\because$ only one pending request is allowed). But y-resp is the last message causing child transition, so child is still at y.

$\Rightarrow$ x-req can not be sent by child since x $<$ y.

$\Rightarrow \Leftarrow$

\end{proof}

\begin{lemma}
Child-state is at x due to parent receiving x-resp and child is at y due to child sending y-resp $\Rightarrow$ y $\le$ x.
\label{CSendPRecv<=}
\end{lemma}

\begin{proof}
Proof by contradiction, assuming x $<$ y.

\begin{figure}
\centering
%\includegraphics{CSendPRecv<=.1}
\end{figure}

Child sends y-resp only after x-resp ($\because$ y-resp is the last message causing child transition).

Child must be in z $>$ y to send y-resp.

$\Rightarrow$ x $<$ z ($\because$ x $<$ y).

$\Rightarrow$ Child must have sent z-req after sending x-resp and transitioning to x, and eventually got back z-resp.

$\Rightarrow$ Parent must have received z-req and sent z-resp, while child-state transitions to z. But this is not possible as x-resp was the last message causing child state transition.

$\Rightarrow \Leftarrow$

\end{proof}

\begin{lemma}
Child-state is at x due to parent sending x-resp and child is at y due to child sending y-resp $\Rightarrow$ y $\le$ x.
\label{CSendPSend<=}
\end{lemma}

\begin{proof}

Proof by contradiction, assuming x $<$ y.

\begin{enumerate}
\item Parent sends x-resp before receiving y-resp and x-resp reaches child after child sends y-resp - Not possible by Lemma \ref{InsaneNetwork}.
\item Parent sends x-resp before receiving y-resp and x-resp reaches child before child sends y-resp.

\begin{figure}
\centering
%\includegraphics{CSendPSend<=.1}
\end{figure}

Child must be in z $>$ y to send y-resp.

$\Rightarrow$ x $<$ z ($\because$ x $<$ y and y $<$ z)

$\Rightarrow$ After child received x-resp, somehow it must reach higher state z. This is possible only if parent sends z-resp, thereby child-state transitioning to z. But x-resp was the last message causing child-state transition.

$\Rightarrow \Leftarrow$

\item Parent receives y-resp before parent sends x-resp.

  \begin{enumerate}
  \item Child sends x-req corresponding to x-resp after sending y-resp.

  \begin{figure}
  \centering
  %\includegraphics{CSendPSend<=.2}
  \end{figure}

  Child is at y $>$ x as y-resp is the last message causing transition for child. Since x $<$ y, child wont send x-req.

  $\Rightarrow \Leftarrow$

  \item Child sends x-req corresponding to x-resp before sending y-resp.

  \begin{figure}
  \centering
  %\includegraphics{CSendPSend<=.3}
  \end{figure}

  Child is at z $<$ x when x-req is sent and z' $>$ y when y-resp is sent.

  $\Rightarrow$ z $<$ z' ($\because$ x $<$ y and z' $>$ y).

  $\Rightarrow$ z'-req must be sent by child before x-resp is received (for x-req). But this is not possible as only one pending request is allowed.

  $\Rightarrow \Leftarrow$

  \end{enumerate}

\end{enumerate}

\end{proof}

\begin{lemma}
Child-state is at x due to parent sending x-resp and child is at y due to child receiving y-resp and parent just received z-req $\Rightarrow$ y = x.
\label{CRecvPSendReq}
\end{lemma}

\begin{proof}

Proof by contradiction, assuming x $\neq$ y.

\begin{figure}
\centering
%\includegraphics{CRecvPSendReq.1}
\end{figure}

Parent sends x-resp only after y-resp ($\because$ x-resp is the last message causing child-state transition).

Child sends corresponding x-req only after receiving y-resp ($\because$ only one pending request is allowed).

$\Rightarrow$ Child can not send z-req before receiving x-resp ($\because$ only one pending request is allowed and x-resp has not reached child yet).

$\Rightarrow \Leftarrow$

\end{proof}

\begin{lemma}
Child-state is at x due to parent receiving x-resp and child is at y due to child sending y-resp and parent just received z-req $\Rightarrow$ y = x.
\label{CSendPRecvReq}
\end{lemma}

\begin{proof}

Proof by contradiction, assuming x $\neq$ y.

\begin{figure}
\centering
%\includegraphics{CSendPRecvReq.1}
\end{figure}

Child sends y-resp only after sending x-resp ($\because$ y-resp is the last message causing child transition).

Parent receives z-req before receiving y-resp ($\because$ x-resp is the last message causing child-state transition).

$\Rightarrow$ Child sent z-req before sending y-resp ($\because$ FIFO ordering between child to parent req, resp).

$\Rightarrow$ y-resp is not voluntary ($\because$ child has pending z-req).

$\Rightarrow$ Parent does not receive z-req till it receives y-resp for the corresponding y-req that parent sent.

$\Rightarrow \Leftarrow$

\end{proof}

\begin{lemma}
Child-state is at x due to parent sending x-resp and child is at y due to child sending y-resp and parent just received z-req $\Rightarrow$ y = x.
\label{CSendPSendReq}
\end{lemma}

\begin{proof}

Proof by contradiction, assuming x $\neq$ y.

\begin{enumerate}
\item Parent sends x-resp before receiving y-resp and x-resp reaches child after child sends y-resp - Not possible because of Lemma \ref{InsaneNetwork}.
\item Parent sends x-resp before receiving y-resp and x-resp reaches child before child sends y-resp.

\begin{figure}
\centering
%\includegraphics{CSendPSend.1}
\end{figure}

y-resp can not be sent voluntarily by child as z-req is pending.

$\Rightarrow$ There is a pending y-req from parent.

$\Rightarrow$ Parent wont receive z-req ($\because$ only one pending request is allowed)

$\Rightarrow \Leftarrow$

\item Parent receives y-resp before parent sends x-resp.

\begin{figure}
\centering
%\includegraphics{CSendPSend.2}
\end{figure}

$\Rightarrow$ There is a pending x-req from child. This is not possible since there are two pending requests from child as child has not received x-resp at this moment.

$\Rightarrow \Leftarrow$

\end{enumerate}

\end{proof}

\begin{lemma}
Child-state is at x due to parent sending x-resp and child was at y before sending z-resp due to child receiving y-resp and parent just received z-resp $\Rightarrow$ y = x.
\label{CRecvPSendResp}
\end{lemma}

\begin{proof}

Proof by contradiction, assuming x $\neq$ y.

\begin{figure}
\centering
%\includegraphics{CRecvPSendResp.1}
\end{figure}

Parent sends x-resp only after y-resp ($\because$ x-resp is the last message causing child-state transition).

Child sends corresponding x-req only after receiving y-resp ($\because$ only one pending request is allowed).

$\Rightarrow$ z-resp is not voluntary.

$\Rightarrow$ Corresponding z-req must be received by child before receiving x-resp ($\because$ x-resp has not yet been received by child when it sent z-resp).

$\Rightarrow$ x-resp is sent by parent between sending z-req ($\because$ FIFO ordering between parent to child messages) and receiving z-resp, which is not possible.

$\Rightarrow \Leftarrow$

\end{proof}

\begin{lemma}
Child-state is at x due to parent receiving x-resp and child is at y before sending z-resp due to child sending y-resp and parent just received z-resp $\Rightarrow$ y = x.
\label{CSendPRecvResp}
\end{lemma}

\begin{proof}

Parent must receive y-resp before receiving z-resp ($\because$ FIFO ordering between child to parent resps). This means that x-resp is not the last message inducing child-state transition.

$\Rightarrow \Leftarrow$

\end{proof}

\begin{lemma}
Child-state is at x due to parent sending x-resp and child is at y before sending z-resp due to child sending y-resp and parent just received z-resp $\Rightarrow$ y = x.
\label{CSendPSendResp}
\end{lemma}

\begin{proof}

Proof by contradiction, assuming x $\neq$ y.

\begin{enumerate}
\item Parent sends x-resp before receiving y-resp.

z-resp reaches parent.

\begin{figure}
\centering
%\includegraphics{CSendPSendResp.1}
\end{figure}

$\Rightarrow$ y-resp reaches parent ($\because$ FIFO ordering between child to parent resps), after x-resp is sent which is not possible.

$\Rightarrow \Leftarrow$

\item Parent receives y-resp before parent sends x-resp.

\begin{figure}
\centering
%\includegraphics{CSendPSendResp.2}
\end{figure}

There is a corresponding x-req that parent receives when sending x-resp. Parent receives x-req before z-resp is received.

$\Rightarrow$ x-req was sent by child before z-resp was sent ($\because$ FIFO ordering between child to parent req, resp).

$\Rightarrow$ z-resp is not voluntary.

$\Rightarrow$ Corresponding z-req was received by child before x-resp was received ($\because$ x-resp was not yet received).

$\Rightarrow$ x-resp was sent by parent between sending z-req ($\because$ FIFO ordering between parent to child messages) and z-resp. This is not possible.

$\Rightarrow \Leftarrow$

\end{enumerate}

\end{proof}

\begin{lemma}
At a moment, child-state is x and child is y $\Rightarrow$ y $\le$ x.
\label{<=}
\end{lemma}

%\begin{proof}
%Initially this is true (all states are invalid).
%
%If only child-state has changed, it can only go up from invalid.
%
%If only child has changed to x, then it can change only due to a x-resp from parent ($\because$ invalid $<$ x). This means child-state also changes. So this is not possible.
%
%If both child and child-state have changed, Lemmas \ref{CRecvPRecvNetwork}, \ref{CRecvPSend<=}, \ref{CSendPRecv<=} and \ref{CSendPSend<=}, examines all scenarios.
%\end{proof}

\begin{lemma}
When parent receives z-req, child-state is x and child is y $\Rightarrow$ y = x.
\label{Req=}
\end{lemma}

%\begin{proof}
%Initially this is true (all states are invalid).
%
%If both child and child-state have changed, Lemmas \ref{CRecvPRecvNetwork}, \ref{CRecvPSendReq}, \ref{CSendPRecvReq} and \ref{CSendPSendReq}, examines all scenarios.
%\end{proof}

\begin{lemma}
When parent receives z-resp, child-state is x and child is y $\Rightarrow$ y = x.
\label{Resp=}
\end{lemma}

%\begin{proof}
%Initially this is true (all states are invalid).
%
%If both child and child-state have changed, Lemmas \ref{CRecvPRecvNetwork}, \ref{CRecvPSendResp}, \ref{CSendPRecvResp} and \ref{CSendPSendResp}, examines all scenarios.
%\end{proof}

\end{document}