### 2012-11-04 paper/isca40format/Setup.tex

\section{Abstract directory-based protocol}
\label{sec:setup}

We first present the state
In an directory-based inclusive cache coherence protocol, the directory keeps
information about the states (of the lines) of all the caches that it services.
The directory essentially has a copy of the 
The directory essentially has a copy of the states of the cache lines of each
cache that it services, but using a more compact encoding. For instance, if a
cache has write permission for a line, no other cache would have any permission
-- in this case, storing just the index of the cache having the write permission is enough,
instead of keeping a replica of the state of the line in each cache.  one
cache can have write permission for a cache line, so it's not necessary to 
The directory can be thought of as having a copy of the states of each cache
line in each state.