### 2012-11-14 paper/isca40format/Variations.tex

\section{Implementing variations of basic MSI protocol}
\label{sec:variations}

%\renewcommand{\fourAngle}[6]{\text{$#1\langle#2\rightarrow#3,#4,#5\rightarrow#6\rangle$}}

\newcommand{\FwdReq}[2]{\text{$FwdReq\langle#1,#2\rangle$}}
\newcommand{\FwdAck}{\text{$FwdAck$}}

%\renewcommand{\Req}[5]{\fourAngle{Req}{#1}{#2}{#3}{#4}{#5}}
%\renewcommand{\Resp}[5]{\fourAngle{Resp}{#1}{#2}{#3}{#4}{#5}}

In this section we describe variations of the basic MSI hierarchical
message-passing directory-based protocol and show how our framework can be used
to design and argue their correctness easily.
%The
%arguments are based on the formally proven Invariants of Figure
%\ref{invariants}. The ordering requirements in Figure \ref{ordering} are
%handled at the protocol level by sending the state at which the sender was in
%just before sending the message as discussed in Section \ref{sec:nofifoorder}

\subsection{MESI optimization}
We first describe the MESI variation of the message-passing MSI protocol for a
two-level hierarchy containing multiple L1 caches and directory shared by all
the L1 caches. On receiving a request from a cache $c$ to upgrade $c$'s state
for address $a$ to $S$, the directory sends permission to upgrade the state of
$c$ for address $a$ to ``exclusive'' ($E$), if the directory state of every
other cache for that line is $I$ (and because of our Invariant
\ref{conservative}, the coherence state of every other cache for that line is
also $I$). Once cache $c$ has upgraded to $E$, it can automatically upgrade to
$M$ on receiving a store request from the processor, without sending any
notification to the directory.  However, if some other cache $c'$ sends a
request to the directory for address $a$ to upgrade to state $S$, the directory
has to ensure that the cache $c$ has downgraded its state from $E$ to $S$. If
$c$ remains at $E$ without having upgraded to $M$, then it doesn't have to send
the data back to the memory, while downgrading to $S$.

From the directory's perspective, states $M$ and $E$ are equivalent -- the only
distinction is that the cache would not send back the data while downgrading the
state for an address from $E$. Thus MESI is simply an optimization over the MSI
protocol in which the cache maintains an extra dirty bit for each address, and
sends back the cache line data to the directory during a downgrade only if the
dirty bit is set (the dirty bit can not be set for an address if the
corresponding state is $S$). If the dirty bit is not set during a downgrade, the
cache sends back a message denoting ``data unmodified'' in the place of actual
data. In the hierarchical setting too, MESI is simply an optimization over MSI
in which a child sends data back to its parent during a downgrade only if the
dirty bit is set.

\subsection{Cache-intervention based protocols}
In the protocols discussed so far, caches can directly communicate only with
their immediate children or with its immediate parent. In cache-intervention
based protocols, communication and data transfer can also take place between
nodes other than the immediate children or the immediate parent. Our framework
can be easily extended to include cache-intervention based protocols where
sibling caches, \ie caches that share the same immediate parent, can communicate
with each other. We show 2 protocols with sibling cache-intervention: a) the MSI
protocol with intervention by a cache in the $M$ state and b) the MOSI protocol. 
Note that in both the protocols, the MESI optimization can be implemented
whenever a child has to transfer data to its parent.

We first describe the two protocols using a two-level hierarchy containing
multiple L1 caches and a directory shared by all the L1 caches, and later
describe the extension to a hierarchical system.

\subsubsection{Intervention by a cache in the $M$ state:}
A downgrade request sent from the directory to a cache $c$ for address $a$ can
sometimes be augmented with a forwarding request \FwdReq{fwdDest}{toState},
which directs the cache to send an upgrade response to $fwdDest$ to upgrade
$fwdDest$'s state for address $a$ to $toState$, on the directory's behalf. A
downgrade response from cache $c$ is also augmented with \FwdAck{}, which
acknowledges the directory that the requested forwarding has been fulfilled.
Note that normal downgrade requests from the directory without a forwarding
request and normal responses from the cache without \FwdAck{} are also sent.

When a directory receives a request from cache $c'$ for address $a$ to upgrade
to state $x$ ($x$ = $M$ or $S$), it determines if some other cache is in
state $M$ for address $a$. If no cache is in state $M$ for address $a$, then the
directory behaves in the same manner as MSI -- we will not discuss this here
again.  But if some cache $c$ is in state $M$ for address $a$, it sends a
request to cache $c$ to downgrade its state for address $a$ to $toCompatible(x)$
(either $I$ or $S$ for $x$ = $M$ or $S$, respectively) augmented with a
forwarding request \FwdReq{c'}{x}. This forwarding request is essentially a
response sent by the directory to cache $c'$ to upgrade the state of $c'$ for
address $a$ to $x$, but it is sent via the cache $c$. The directory waits for a
downgrade response along with a \FwdAck{} from $c$ before processing further
requests for $a$.

If cache $c$ receives the downgrade request from the directory along with the
forward request \FwdReq{c'}{x}, while it is still in state $M$ for address $a$,
$c$ will send a response to $c'$ authorizing the upgrade of $state[c'][a]$ to
$x$, along with the data associated with address $a$. It will also send a
downgrade response to the directory notifying that $c$ has downgraded state $a$
to $toCompatible(x)$ along with a \FwdAck{}; it doesn't transfer the data to the
directory/memory. However, if cache $c$ has already downgraded to a lower state
than $M$ for address $a$ (because of eviction due to replacement), then the
cache simply drops the request just like in the MSI protocol discussed earlier;
when $c$ downgraded its state from $M$ for address $a$, it must have sent a
downgrade response to the directory for address $a$ along with the data (subject
to MESI optimization), just like in the MSI protocol discussed earlier,
albeit without a \FwdAck{}.

If the directory receives a downgrade response for address $a$ from cache $c$
along with a \FwdAck{}, the directory knows that $c$ has sent a response to cache
$c'$ authorizing $c'$ to upgrade its state for $a$ to $x$. But if the directory
receives a downgrade response for address $a$ from cache $c$ without a \FwdAck{},
the directory knows that $c$ has not sent a response to cache $c'$. In this
case, the directory sends a response along with the data to $c'$ authorizing it
to upgrade its state for address $a$ to $x$. Figure \ref{msi-inter} shows the
order of messages transferred between the caches in the scenario when forwarding
happens as well the in the scenario when no forwarding happens.

$c'$ would eventually get the upgrade to $x$ response for address $a$ either
from the directory or the cache $c$ -- it can not distinguish between the two
cases.

The protocol given above satisfies all the constraints \ref{} to \ref{}, except,
potentially, Constraint \ref{orderreq}. Before the upgrade response for address
$a$ sent from $c$ to $c'$ has reached $c'$, $c'$ could have received a downgrade
request sent from the parent. This happens as follows (Figure
\ref{msi-fifo-needed}, with the numbers indicating the order in which the
messages are sent, for outgoing edges, and received for incoming edges). Cache
$c$ sends the upgrade response (and data) to cache $c'$, and its own downgrade
response along with the \FwdAck{} to the directory. The directory receives the
downgrade response from $c$. Soon afterwards, the directory wants to downgrade
the state of $a$ in $c'$ to $I$ (let's say, because some other cache wants to
upgrade its state for $a$ to $M$).  The directory sends a downgrade to $I$
request to $c'$ for address $a$. This downgrade request to $c'$ from the parent
reaches $c'$ before the (forwarded) upgrade response from $c$, violating
Constraint \ref{orderreq}. However, as we saw in Section \ref{sec:nofifoorder},
Constraint \ref{orderreq} can be satisfied by tagging the downgrade request with
the state the directory was in for cache $c'$ and address $a$ right before the
request to $c'$ was sent -- FIFO ordering is not required from the network. $c'$
will not process the downgrade request till it has received the upgrade response
from $c$.

Since the protocol satisfies all the constraints \ref{} to \ref{}, it satisfies
all the invariants in Figure \ref{invariants}. Thus the directory is guaranteed
to have a conservative version of the states of each of the caches for each
address by Invariant \ref{conservative}. When the directory sends a downgrade
request for address $a$ to $c$ tagged with a \FwdReq{c'}{x} (when $dir[c][a]$ is
$M$), the directory assumes that both $c$ and $c'$ are in state $M$. However, it
is guaranteed that $c'$ can not upgrade to $x$ unless $c$ has sent an upgrade
response to $c'$, which guarantees that $c$ would have downgraded to
$toCompatible(x)$. This ensures both the Single-writer Invariant
\ref{singleWriter} as well as Read-from-last-writer Invariant \ref{readLast}.

Note that in our protocol, \FwdAck is sent from the forwarding cache $c$ rather
than the final cache $c'$ which is the case in commonly used protocols
\cite{hammer, gems}. This unblocks the directory to serve further requests for
the same address as soon as \FwdAck is received from $c$, instead of waiting for
an acknowledgement from $c'$.

Thus we have mapped the complex protocol with intervention by a cache in $M$
state to a simpler MSI protocol and proved its correctness using our framework.

\subsubsection{MOSI protocol}