### 2012-11-18 paper/isca40format/ArvindGood/Init.tex

\section{Initialize}
\label{sec:init}

Cache coherence protocols maintain coherence by ensuring the following two
invariants:

\begin{inv}
\textbf{Single-writer Invariant}: When a cache has permission to modify data for
an address, no other cache has read or write permission for that
address.
\label{singleWriter}
\end{inv}

\begin{inv}
\textbf{Read-from-last-writer}: When a cache reads data for an address, it reads
the last modified version of the data for the address. If the data for an
address has not been modified, the cache reads the data from memory.
\label{lastRead}
\end{inv}

Because of the requirement to maintain the Single-writer property (Invariant
\ref{singleWriter}), it makes sense to talk about compatibility of the states of
two caches for an address. Two coherence states $x$ and $x$ are \emph{compatible}
if two caches having the same address in states $x$ and $x$ do not violate the
single-writer property.