### 2012-11-18 paper/isca40format/stuff/AbstractProtocol.tex
stuff

\section{Abstract directory-based protocol}

\label{sec:abstract}

\newcommand{\One}[2]{\text{#1$\langle#2\rangle$}}
\newcommand{\qOne}[2]{\text{#1[$#2$]}}
\newcommand{\qTwo}[3]{\text{#1[$#2$][$#3$}}
\newcommand{\two}[3]{\text{#1$\langle#2,#3\rangle$}}
\newcommand{\three}[4]{\text{#1$\langle#2,#3,#4\rangle$}}
\newcommand{\four}[5]{\text{#1$\langle#2,#3,#4,#5\rangle$}}
\newcommand{\True}{\texttt{True}}
\newcommand{\False}{\texttt{False}}
\newcommand{\assigns}{:=}
\newcommand{\AndAnd}{\;\&\&\;}
\newcommand{\OrOr}{\;||\;}
\newcommand{\Inv}{\text{$Inv$}\;}

\definecolor{trig-color}{gray}{.4}

We will give the abstract directory-based protocol for a hierarchical system of
caches. The caches are assumed to be organized in the form of a tree. The leaves
of this tree represent the L1 caches, with each L1 cache connected to exactly
one core. The root of the tree is a single Last-level cache (LLC), connected to
the main memory.

A node keeps information about the coherence state of each cache line in the
node as well as in each of its children, \ie the directory.
Figure~\ref{table:states} shows the precise information that a node keeps for
each cache line in the node. The notation we adopt is primarily for conciseness
of description. For instance, the directory need not contain the exact copy of
each of the child's coherence state for each cache line; instead the encoding
can be more efficient. All the information (except $present, data, state$ and
$dir$) corresponds to a request being handled for a given address; an
implementation can store this information only for the requests being processed
instead of for each address.

\begin{figure}
\begin{tabular}{|l|p{.64\columnwidth}|}
\hline
isPresent[a] & Is cache line for address $a$ present?\\
\hline
data[a] & Cache line data for address $a$\\
\hline
state[a] & Coherence state for address $a$\\
\hline
dir[a][c] & Directory's copy of the coherence state for address $a$ in child
$c$\\
\hline
isCWait[a] & Is node waiting for a response from any child for address $a$?\\
\hline
cWaitState[a] & The coherence state to which the node has requested the children to
downgrade to, for address $a$\\
\hline
isPWait[a] & Is node waiting for a response from its parent for
address $a$?\\
\hline
isPReq[a] & Is node handling a request from parent for addess $a$\\
\hline
isCReq[a] & Is node handling a request from a child for address $a$\\
\hline
whichCReq[a] & The child whose request the node is handling for address $a$\\
\hline
isEvict[a] & Is the cache line for address $a$ being evicted?\\
\hline
replacingAddr[a] & If the cache line for address $a$ is being evicted, the address replacing the line\\
\hline
\end{tabular}
\caption{Information that the node keeps for a cache line}
\label{table:states}
\end{figure}

The values of the coherence state (and hence the directory state) belong to a
totally ordered set, \ie a set in which a \emph{less than} ($<$) relation is
defined between any two set elements and this relation is
transitive. For instance, in an MSI protocol, the values of the coherence state
are from the set $\{M, S, I\}$ where $I < S$, $S < M$ and $I < M$. Our abstract
protocol does not assign any semantics or meaning to the values of the coherence
states. In a concrete cache coherence protocol, the values of the states are
associated with permissions for reading and writing, and the state with fewer
permissions is $<$ the state with more permissions. We use \Inv to denote the
least element in the set containing the values of the coherence state of a line.
Typically, this denotes that the node has no permissions for reading or writing
that line.

In order to increase the coherence state of a cache line, a node has to send an
\emph{upgrade request} to its parent and may eventually get an \emph{upgrade
response} from the parent. In order to decrease the coherence state of a cache
line for one of the node's children, the node has to send a \emph{downgrade
request} to that child, and may eventually get a \emph{downgrade response} from
that child.

Figure~\ref{table:messages} shows the incoming and outgoing messages to and from
the node, respectively.

\begin{figure}
\begin{tabular}{|l|p{.6\columnwidth}|}
\hline
\multicolumn{2}{|c|}{Incoming messages}\\
\hline
\three{ReqFromC}{c}{a}{x} & Request from child $c$ upgrade $c$'s coherence
state for address $a$ to $x$\\
\two{ReqFromP}{a}{x} & Request from parent to downgrade node's coherence
state for address $a$ to $x$\\
\three{RespFromC}{c}{a}{x} & Message from child $c$ denoting that $c$'s
coherence state for address $a$ has downgraded to $x$\\
\three{DataFromC}{c}{a}{d} & Data $d$ of the cache line received from child $c$
for address $a$\\
\two{RespFromP}{a}{x} & Response from parent giving permission to
upgrade the node's coherence state for address $a$ to $x$\\
\two{DataFromP}{a}{d} & Data $d$ of the cache line received from parent for
address $a$\\
\hline
\multicolumn{2}{|c|}{Outgoing messages}\\
\hline
\three{ReqToC}{c}{a}{x} & Request to child $c$ to downgrade $c$'s
coherence state for address $a$ to $x$\\
\two{ReqToP}{a}{x} & Request to parent to upgrade node's coherence
state for address $a$ to $x$\\
\three{RespToC}{c}{a}{x} & Response to child $c$ granting permission to
upgrade $c'$ coherence state for address $a$ to $x$\\
\three{DataToC}{c}{a}{d} & Data $d$ of the cache line sent to the child
$c$ for address $a$\\
\two{RespToP}{a}{x} & Message to parent denoting that the node's
coherence state for address $a$ has downgraded to $x$\\
\two{DataToP}{c}{d} & Data $d$ of the cache line sent to the parent
for address $a$\\
\hline
\end{tabular}
\caption{Incoming and outgoing messages}
\label{table:messages}
\end{figure}

Each message coming from a particular child or the parent is independent of the
messages coming from another source, in addition to being independent of the
messages for a different address. The node can process messages coming from
different sources, or corresponding to different addresses, in any order,
subject to the ordering requirements in Figure~\ref{list:messageOrder}. The
message ordering restrictions are necessary to avoid deadlocks. We later show
that is easy to enforce this ordering even if the underlying network doesn't
enforce point-to-point FIFO ordering or priorities between different kinds of
messages.

\floatstyle{boxed} 
\restylefloat{figure}

\begin{figure}
\begin{minipage}{\columnwidth}
\begin{enumerate}
\item A request or a response should not be handled before handling a response
sent earlier from the same source for the same address

\item A request from a child for an address should not block or prevent the
handling of a request from the parent for the same address

\item A request for an address should not block or prevent the handling of a
response from any source for any address
\end{enumerate}
\end{minipage}
\caption{Message ordering, and blocking restrictions}
\label{list:messageOrder}
\end{figure}

\floatstyle{plain} 
\restylefloat{figure}

%To allow maximum flexibility of implementation, the node is given separate
%access queues for each kind of message (request of response) from each source
%and for each address (\ie \qTwo{ReqFromC}{c}{a}, \qOne{ReqFromPQ}{a},
%\qTwo{RespFromCQ}{c}{a}, \qTwo{ReqToCQ}{c}{a}, \qOne{RespToPQ}{a}, \etc). An
%implementation of a protocol using our methodology does not have to keep a
%separate access queue for each source and each address, but it must adhere to
%the blocking restrictions in Figure~\ref{list:messageOrder}.  These
%restrictions are necessary to avoid deadlocks. The flexibility of the abstract
%protocol allows one to design a wide range of blocking policies.

Our abstract protocol also depends on the designer correctly defining the three
functions shown in Figure~\ref{table:functions}. These functions abstract away
the coherence aspect of the cache coherence protocol. Note that $maxCompat$
function can be derived from $isCompat$ function, if $isCompat$ is defined
correctly. Figure~\ref{fig:msiFunc} shows the functions defined for MSI (``--''
denotes undefined).

\begin{figure}
\begin{tabular}{|c|p{.63\columnwidth}|}
\hline
isCompat(x, y) & Returns if two different sibling caches can have coherence
states $x$ and $y$, respectively, for the same address or not\\
\hline
maxCompat(x) & Returns a coherence state value $y$ such that, $\forall z \le y,
isCompat(x, z)$ and $\forall w > y, !isCompat(x, w)$\\
\hline
dataXfer(x, y) & Return if cache line data has to be transferred by a cache to its
parent on downgrading ist coherence state from $x$ to $y$ or not\\
\hline
\end{tabular}
\caption{Functions to abstract away the coherence aspect}
\label{table:functions}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}{.45\columnwidth}
\centering
\begin{tabular}{|c||c|c|c|}
\hline
& I & S & M\\
\hline\hline
I & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\hline
S & $\checkmark$ & $\checkmark$ & $\times$\\
\hline
M & $\checkmark$ & $\times$ & $\times$\\
\hline
\end{tabular}
\label{table:isCompat}
\caption{$isCompat$ for MSI}
\end{subfigure}
\quad
\begin{subfigure}{.45\columnwidth}
\centering
\begin{tabular}{|c||c|c|c|}
\hline
\backslashbox{$x$}{$y$} & I & S & M\\
\hline\hline
I & -- & -- & --\\
\hline
S & $\times$ & -- & --\\
\hline
M & $\checkmark$ & $\checkmark$ & --\\
\hline
\end{tabular}
\label{table:dataXfer}
\caption{$dataXfer(x, y)$ for MSI}
\end{subfigure}
\caption{Coherence abstraction for MSI}
\label{fig:msiFunc}
\end{figure}

We now describe the how a node handles requests and responses.

Note
that a message handler would be triggered only if the corresponding message is
present. If the required resources for handling a message are not present (for
example, if the handling of the request requires sending further requests to
its children or the parent, and the corresponding send buffers are full), then
the handler for that message would not be triggered.

Figure~\ref{alg:handlePReq} specifies how a node handles an incoming request
from the parent for downgrading the coherence state of address $a$ to $x$. Note
that a request from the parent can be handled successfully only if the node is
not already waiting for a response from some child.

Figure~\ref{alg:handleCReq} specifies how a node handles an incoming request
from a child for upgrading the child's coherence state for address $a$ to $x$.
Note that a request from a child can be handled successfully only if the node
is not already waiting for a response from the parent or from some child.

Figure~\ref{alg:handlePResp} specifies how a node handles an incoming response
from the parent for permitting upgrade of the coherence state of address $a$ to
$x$.

Figure~\ref{alg:handleCResp} specifies how a node handles an incoming response
from a child notifying the downgrade of the child's coherence state for address
$a$ to $x$.

XXX Figure~\ref{alg:handleEvict} specifies how a cache line replacement is handled.

\floatstyle{boxed} 
\restylefloat{figure}

\begin{figure}
\begin{algorithmic}
\Rule{handlePReq}{a}
   \State let mesg \two{ReqFromP}{a}{x} exist in \qOne{ReqFromPQ}{a}
   \If {$\forall c, !isCWait(c][a)$}
     \If {$state(a) \le x$}
        \State remove$(\two{ReqFromP}{a}{x})$
     \State remove$(\two{ReqFromP}{a}{x})$
     \ElsIf {$\forall c, dir(c][a) \le x$}
        \State remove$(\two{ReqFromP}{a}{x})$
        \If {dataXfer$(state(a), x)$}
           \State send$(\two{DataToP}{a}{data(a)})$
        \EndIf
        \State $state(a) \assigns x$
        \State send$(\two{RespToP}{a}{x})$
     \Else
        \ForAll{$c$}
           \If {$dir(c][a) > x$}
              \State $isCWait(c][a) \assigns \True$
              \State $cWait(c][a) \assigns x$
              \State send$(\three{ReqToC}{c}{a}{x})$
           \EndIf
        \EndFor
        \State \textcolor{trig-color}{$whoTrigger(a) \assigns Parent$}
     \EndIf
   \EndIf
\EndRule
\end{algorithmic}
\caption{Handle request from parent}
\label{alg:handlePReq}
\end{figure}

\begin{figure}
\begin{algorithmic}
\Rule{handleCReq}{$c, a$}
   \State let mesg \two{ReqFromC}{c}{a}{x} exist in \qTwo{ReqFromCQ}{c}{a}
   \If {$isPresent(a)$}
      \If {$!isPWait(a) \AndAnd \forall c, !isCWait(c][a)$}
         \If {$state(a) \ge x \AndAnd$ \\
               \hspace{2.2cm} $\forall c' \neq c, dir(c', a) \le \text{maxCompat}(x)$}
            \State remove$(\three{ReqFromC}{c}{a}{x})$
            \If {$dir(c,a) \neq \Inv)$}
               \State send$(\three{DataToC}{c}{a}{data(a)})$
            \EndIf
            \State $dir(c][a) \assigns x$
            \State send$(\three{RespToC}{c}{a}{x})$
         \Else
            \If {$state(a) < x$}
               \State $isPWait(a) \assigns \True$
               \State send$(\two{ReqToP}{a}{x})$
            \EndIf
            \ForAll {$c' \neq c$}
               \If {$dir(c', a) > \text{maxCompat}(x)$}
                  \State $isCWait(c', a) \assigns \True$
                  \State $cWait(c', a) \assigns \text{maxCompat}(x)$
                  \State send$(\three{ReqToC}{c'}{a}{\text{maxCompat}(x)})$
               \EndIf
            \EndFor
            \State \textcolor{trig-color}{$whoTrigger(a) \assigns Child$}
            \State \textcolor{trig-color}{$childTrigger(a) \assigns c$}
         \EndIf
      \EndIf
   \Else
      \State \textbf{let} $a' \assigns \text{getReplace}(a)$
      \If {$!isPWait(a') \AndAnd \forall c', !isCWait(c', a')$}
         \State \textcolor{trig-color}{$whoTrigger(a) \assigns Evict$}
         \State \textcolor{trig-color}{$childTrigger(a) \assigns c$}
         \State $replacingAddr(a') \assigns a$
         \State \textcolor{trig-color}{\textbf{trigger} handleEvict($a$)}
      \EndIf
   \EndIf
\EndRule
\end{algorithmic}
\caption{Handle request from child}
\label{alg:handleCReq}
\end{figure}

\begin{figure}
\begin{algorithmic}
\Rule{handleEvict}{$a$}
   \If {$!isPresent(a) \OrOr \forall c, dir(c][a) = \Inv$}
      \If {$state(a) \neq \Inv$}
         \State send$(\two{RespToP}{a}{\Inv})$
         \If {dataXfer$(state(a), \Inv)$}
            \State send$(\two{DataToP}{a}{data(a)})$
         \EndIf
      \EndIf
      \State $isPresent(a) \assigns \False$
      \State allocLine$(replacingAddr(a))$
      \State \textcolor{trig-color}{$whoTrigger(a) \assigns None$}
      \State \textcolor{trig-color}{\textbf{trigger} handleCReq($childTrigger(a), a$)}
   \Else
      \If {$\forall c, !isCWait(c][a)$}
         \ForAll{$c$}
            \If {$dir(c][a) > \Inv$}
               \State $isCWait(c][a) \assigns \True$
               \State $cWait(c][a) \assigns \Inv$
               \State send$(\three{ReqToC}{c}{a}{\Inv})$
            \EndIf
         \EndFor
      \EndIf
   \EndIf
\EndRule
\end{algorithmic}
\caption{Evict a line in address $a$}
\label{alg:handleEvict}
\end{figure}

\begin{figure}
\begin{algorithmic}
\Rule{handlePResp}{$a$}
   \State let mesg \two{RespFromP}{a}{x} exist in \qOne{RespFromPQ}{a}
   \State remove$(\two{RespFromP}{a}{x})$
   \If {$state(a) \neq \Inv$}
      \State let mesg \two{DataFromP}{a}{d} exist in \qOne{DataFromPQ}{a}
      \State remove$(\two{DataFromP}{a}{d})$
      \State $data(a) = d$
   \EndIf
   \State $state(a) \assigns x$
   \State $isPWait(a) \assigns \False$ 
   \State \textcolor{trig-color}{\textbf{trigger} handleCReq($childTrigger(a), a$)}
\EndRule
\end{algorithmic}
\caption{Handle response from parent}
\label{alg:handlePResp}
\end{figure}

\begin{figure}
\begin{algorithmic}
\Rule{handleCResp}{$c, a$}
   \State let mesg \three{RespFromC}{c}{a}{x} exist in \qTwo{RespFromCQ}{c}{a}
   \State remove$(\three{RespFromC}{c}{a}{x})$
   \If {dataXfer$(dir(c,a), x)$}
      \State let mesg \three{DataFromC}{c}{a}{d} exist in \qTwo{DataFromPQ}{c}{a}
      \State remove$(\three{DataFromC}{c}{a}{d})$
      \State $data(a) = d$
   \EndIf
   \State $dir(c,a) \assigns x$
   \If {$isCWait(c][a) \AndAnd cWait(c][a) >= x$}
     \State $isCWait(c][a) \assigns \False$
   \EndIf
   {\color{trig-color}
   \If {$whoTrigger(a) = Parent$}
      \State \textbf{trigger} handlePReq($a$)
   \ElsIf {$whoTrigger(a) = Evict$}
      \State \textbf{trigger} handleEvict($(a$)
   \ElsIf {$whoTrigger(a) = Child$}
      \State \textbf{trigger} handleCReq($childTrigger(a), a$)
   \EndIf}
\EndRule
\end{algorithmic}
\caption{Handle response from child}
\label{alg:handleCResp}
\end{figure}

Function \emph{getReplace(a)}, used in \textsc{handleCReq} gives the address of
the line that can be replaced in order to accommodate the line in address $a$.
Note that if the line selected for eviction is waiting for a response from the
parent or from any child, then it can not be replaced.

Figure~\ref{alg:allocLine} shows the initiation that has to happen for
allocating a new cache line for an address.

\begin{figure}
\begin{algorithmic}
\Procedure{allocLine}{$a$}
  \State $isPresent(a) \assigns \True$
  \State \textcolor{trig-color}{$whoTrigger(a) \assigns None$}
  \State $state(a) \assigns \Inv$
  \ForAll {$c$}
     \State $dir(c][a) \assigns \Inv$
     \State $isCWait(c][a) \assigns \False$
  \EndFor
  \State $isPWait(a) \assigns \False$
\EndProcedure
\end{algorithmic}
\caption{Allocate a new cache line for address $a$}
\label{alg:allocLine}
\end{figure}

Each handler will be triggered repeatedly till the corresponding message has
been removed. In order to avoid starvation, the following two conditions must be
maintained:

\begin{enumerate}

\item Once a response is available from any source, the handler for that
response must be triggered eventually.

\item After handling a response (or an eviction) the corresponding handler
must trigger the original handler that requested the received response (or made
an eviction request). The $whoTrigger(a)$ keeps track of the original handler
for each address $a$. In case of a $Child$ trigger, the actual child
corresponding to the original handler is kept track of in $childTrigger(a)$, for
address $a$, in which case $handleCReq$ is triggered for the outstanding request
from that child. In the figures, the pseudo-code corresponding to this is grayed out

\end{enumerate}

The L1 caches, \ie the leaf nodes, do not have any children. Instead, they
handle requests from cores. A store word request is essentially a request by
the core for write permissions for the corresponding cache line, and a load
word request is a request for read permissions. These permissions can be mapped
to the values of the coherence state. Because of the absence of children any
request from the parent as well as any eviction can be handled without
producing more requests. As a result, we do not need a separate handler for
eviction requests. We also do not need to store $whoTrigger(a)$, since the only
trigger that produces more requests, and hence has to wait for response is a
request from the core. We omit the pseudo-code for handling a request from the
core because it is very similar to an instance of \textsc{handleCReq} with no
children.

%Figure~\ref{alg:handleCoreReq} shows the handler for a request from the core,
%which is very similar to a combination of \texttt{handleCReq} and
%\texttt{handleEvict}. The code shows that the cache line is sent back to the
%core for any request, which is not necessary for a store request in an actual
%implementation.
%
%\begin{figure}
%\begin{algorithmic}
%\Rule{handleCoreReq}{\three{ReqFromCore}{a}{x}}
%  \If {$isPresent(a)$}
%     \If {$!isPWait(a)$}
%        \If {$state(a) \ge x$}
%           \State remove$(\three{ReqFromCore}{a}{x})$
%           \State send$(\three{RespToCore}{a}{x}{data(a)})$
%        \ElsIf {$state(a) < x$}
%              \State $isPWait(a) \assigns \True$
%              \State send$(\two{ReqToP}{a}{x})$
%        \EndIf
%     \EndIf
%  \Else
%     \State \textbf{let} $a' \assigns \text{getReplace}(a)$
%     \If {$!isPWait(a')$}
%        \If {$state(a') \neq Inv$}
%           \State send($\two{RespToP}{a'}{Inv}$)
%           \If {dataXfer($state(a'), Inv$)}
%              \State send($\two{DataToP}{a'}{Inv}$)
%           \EndIf
%        \EndIf
%        \State $isPresesnt(a') \assigns \False$
%        \State allocLine($a$)
%        \State $isPWait(a) \assigns \True$
%        \State send$(\two{ReqToP}{a}{x})$
%     \EndIf
%  \EndIf
%\EndRule
%\end{algorithmic}
%\caption{Handle request from core}
%\label{alg:handleCoreReq}
%\end{figure}

\floatstyle{plain} 
\restylefloat{figure}

%The following are the messages transferred to and from the node
%For the abstract protocol, we consider a hierarchical system of caches. The
%caches are organized in the form of a tree. The leaves of this tree are the L1
%caches connected to each core, and the root of the tree is the LLC, connected to
%the main memory. Each node (leaf and non-leaf) contains several cache lines.
%For each cache line in a node, there is a tag and the coherence state of the
%line. Non-leaf nodes contain the directory of each line in addition to the
%coherence state. Each node can communicate only with its parent or children
%nodes.
%
%The directory of a cache line in a node has a copy of the coherence state of
%that cache line in each child-node. Note that the directory can use a clever bit
%encoding in order to reduce the number of bits used, but the information content
%of the directory remains the same. So, for the purposes of designing and
%verification of a protocol, our assumption about the directory state containing
%a copy of each cache state for each line is valid; only the bit counts in the
%implementation varies.
%
%The abstract protocol defines the behavior of each node, \ie the transitions of
%the directory as well as the coherence state on sending or receiving a message.
%The transitions are based on local information, \ie a node can access or modify
%only the states present in the node.
%
%The values of the coherence state (and hence the directory state) belong to a
%totally ordered set, \ie a set in which a \emph{less than} ($<$) relation is
%defined between every pair of elements in that set.
%
%We will use the following notation to specify the information associated with a
%particular cache line for address $a$.
%
%\begin{itemize}
%
%\item $data[a]$ denotes the data for address $a$.
%
%\item $state[a]$ denotes the coherence state of address $a$.
%
%\item $dir[a](c)$ denotes the directory's copy of the coherence state of address
%$a$ in cache $c$.
%
%\item $tag[a]$ denotes the tag of address $a$.
%
%\item $isCWait[a]$ denotes if the node is waiting for a response from any of its
%child regarding address $a$.
%
%\item $cWait[a]$ denotes the request that a node has sent its children for
%address $a$ and hence waiting for a response.
%
%\item $isPWait[a]$ denotes if the node is waiting for a response from its parent
%regarding address $a$.
%
%\item $pWait[a]$ denotes the request that a node has sent its parent for address
%$a$ and hence waiting for a response.
%
%\end{itemize}
%
%Each node sends requests and responses to its children and its parent to
%orchestrate changing of coherence states. Requests are sent to a node's parent
%for upgrading the coherence state of a line in that node, and to the node's
%children to downgrade the coherence states of the corresponding children.
%Responses received from a parent signify that the directory in the parent has
%been upgraded appropriately for that cache. Responses received from a child
%signify that the coherence state of the child has been downgraded appropriately.
%Each message coming from a particular child or the parent is independent of the
%messages coming from another source, in addition to being independent of the
%messages for a different address. The node can process messages coming from
%different sources, and for different addresses in any order. From the same
%source and for the same address, we assume that neither a response not a request
%can not overtake an earlier response
%
%Even though our abstract protocol requires these message orderings, we later
%show that is easy to enforce these even if the underlying network doesn't
%enforce point-to-point FIFO ordering or priorities between different kinds of
%messages.
%
%To allow maximum flexibility of implementation, the node has separate access
%queues for each kind of message (request of response) from each source and for
%each address. An implementation of a protocol using our methodology does not
%have to keep a separate access queue for each source and each address, but it
%has to adhere to the following rules: a) a request from any child should not
%cause head-of-line blocking of a request from a parent, and b) no request should
%cause head-of-line blocking of any response. As we show later, these rules are
%necessary to avoid deadlocks. This flexibility allows one to use a wide range of
%blocking policy.
%
%The following notation will be used for the requests and responses sent and
%received:
%
%\begin{itemize}
%
%\item $p.sendReq(a, to)$ denotes sending a request to the parent
%in order to upgrade the coherence state of address $a$ in the node to state
%$to$.
%
%\item $c.sendReq(a, to)$ denotes sending a request to child $c$ in order to
%downgrade the coherence state of address $a$ in the $c$ to state $to$.
%
%\item $p.sendResp(a, to, d)$ denotes sending a response message to the parent
%signifying that the coherence state of the node for address $a$ has changed to
%$to$. $d$ denotes the cache line sent (if necessary) by the node to the parent.
%
%\item $c.sendResp(a, to, d)$ denotes sending a response message to child $c$
%signifying that the directory state of the node for address $a$ and child $c$
%has changed to $to$. $d$ denotes the cache line sent (if necessary) by the node
%to child $c$.
%
%\item $p.peekReq[a]$ denotes the request from the parent for address $a$. The
%message contains the value of the coherence state for address $a$ that this node
%should downgrade to.
%
%\item $c.peekReq[a]$ denotes the request from child $c$ for address $a$. The
%message contains the value of the coherence state for address $a$ that child $c$
%wants to upgrade to.
%
%\item $p.deqReq[a]$ denotes finishing up of the processing of the request from
%the parent for address $a$ and hence removing the request.
%
%\item $c.deqReq[a]$ denotes finishing up of the processing of the request from
%child $c$ for address $a$ and hence removing the request.
%
%\item $p.peekResp[a]$ denotes the response from the parent for address $a$. The
%message contains a pair $\langle to, d\rangle$, where $to$ denotes the value of
%the coherence state for address $a$ that this node should upgrade to, and $d$
%represents the cache line data sent by the parent (if necessary)
%
%\item $c.peekResp[a]$ denotes the response from child $c$ for address $a$. The
%message contains a pair $\langle to, d\rangle$, where $to$ denotes the value of
%the coherence state for address $a$ that child $c$ has downgraded to, and $d$
%represents the cache line data sent by the child (if necessary)
%
%\item $p.deqReq[a]$ denotes finishing up of the processing of the response from
%the parent for address $a$ and hence removing the response.
%
%\item $c.deqReq[a]$ denotes finishing up of the processing of the response from
%child $c$ for address $a$ and hence removing the response.
%
%\end{itemize}
%