### 2012-11-19 paper/isca40format/Theorems.tex

\section{Theorems}
\label{sec:theorems}

If Invariants \ref{blocking} to \ref{evict} are obeyed, then we have formally
proved that the following theorems hold. The formal proof is beyond the scope of
the paper.

\begin{theorem}
It a thread has to send a message, it will eventually be able to send the
message.
\label{canSend}
\end{theorem}

\begin{theorem}
If a thread sends a request, it will eventually get a response.
\label{willRecv}
\end{theorem}

\begin{theorem}
If $p$ sends \Resp{p}{c}{a}{x} to $c$, where $p = c.parent$, then $p.dir[c][a]$
just before $p$ sends the response is the same as $c.state[a]$ just before $c$
receives the response.
\label{pcSame}
\end{theorem}

\begin{theorem}
If $c$ sends \Resp{c}{p}{a}{x} to $p$, where $p = c.parent$, then $c.state[a]$
just before $c$ sends the response is the same as $p.dir[c][a]$ just before $p$
receives the response.
\label{cpSame}
\end{theorem}

\begin{theorem}
Conservative: For two nodes $p, c$, where $p = c.parent$, $\forall a,
p.dir[c][a] \ge c.state[a]$.
\label{conservative}
\end{theorem}

These theorems are needed to guarantee that Figure \ref{realistic} protocol is
indeed correct. If, for instance, a thread wants to send a message in the output
channel of the node and is not able to send one forever, then the thread can not
proceed further. Similarly, if a thread is waiting for a response (for a request
previously sent), and it never gets the response, the thread can never proceed
further. These scenarios create a deadlock.

We will illustrate the need for Theorems \ref{pcSame}, \ref{cpSame} and
\ref{conservative} using a 2-level of caches having two L1 caches $c_1$ and
$c_2$, and one shared L2 cache $p$.

Theorem \ref{pcSame} and \ref{cpSame} are important to make sure that the data
transfers happen correctly. Let $p.dir[c_1][a] = S$, $p.dir[c_2][a] = I$,
$c_1.state[a] = S$, and $c_2.state[a] = I$. $c_2$ gets a store request, and
sends \Req{c_1}{p}{a}{M} to $p$. $p$ receives the request, but assumes that
$c_1$ already has data for $a$ since $p.dir[c_1][a] = S$. Theorem \ref{pcSame}
is violated. So, it does not transfer the data to $c_1$, resulting in $c_1$
waiting for data, and hence deadlock. A dual condition can be constructed
similarly, in which $c_1$ downgrades its state for address $a$ from $S$ to $I$,
and the parent assumes that $c_1$ is in state $M$, and keeps waiting for data.

Theorem \ref{conservative} is important because a parent node consults its
directory to decide the children it needs to send a downgrade request to.  Let
$p.dir[c_1][a] = S$, $c_1.state[a] = S$, $p.dir[c_2][a] = I$ but
$c_2.state[a] = S$. If $c_1$ gets a store request from the processor and sends
\Req{c_1}{p}{a}{M} to $p$, $p$ will assume that $c_2$'s state for address $a$
is also $I$ and hence sends \Resp{p}{c_1}{a}{M} to $c_1$. But since the real
state of  $c_2$ for address $a$ is $S$, it will break Single-writer Invariant
\ref{singleWriter}.

If Theorems \ref{canSend}, to \ref{cpSame} holds, then the procedures given in
Figure \ref{realistic} can not deadlock because it will always be able to a)
send a request, b) send back a response, c) get back a response for a request it
sent, d) receive data when required. Procedure \uReq{} in Figure
\ref{msi-template} relies on Theorem \ref{conservative} to maintain the
Single-writer Invariant \ref{singleWriter}.