### 2012-11-20 paper/isca40format/Invariants.tex

\section{Local Properties}
\label{sec:localProperties}

In this section we list the local properties that are required to ensure the
invariants in Section \ref{sec:properties}. We can easily see by inspection that the distributed MSI protocol in Figure \ref{realistic} does not violate the local properties in this section.

 which in turn makes the 
distributed cache coherence protocol given in Figure \ref{realistic} correct. We
 give the intuition behind the these invariants whenever applicable, though the
formal proof on why these local properties guarantee the invariants of Section
\ref{sec:theorem} is beyond the scope of this paper.

In all the scenarios that we present below, we will use a system containing two
levels of caches, L1 and L2. There are 2 L1 caches $c_1$ and $c_2$, and both are
connected to the shared L2. Both $c_1$ and $c_2$ can handle multiple outstanding
requests from the cores.

\floatstyle{boxed}
\restylefloat{figure}

\begin{figure}\small
\begin{inv}
An incoming response message for an address $a$ from a child $c$ should not be
received before another incoming response message from the same $c$ and the
same $a$ sent earlier has been received\label{cRespFifo}
\end{inv}
\begin{inv}
An incoming request message for an address $a$ from a source $n$ should not be
received before another incoming response message from the same $n$ and the
same $a$ sent earlier has been received\label{reqNoOvertakeResp}
\end{inv}
\caption{Ordering requirements}
\label{order}
\end{figure}

Let's illustrate the need for Local-Property \ref{cRespFifo}. Let's say $c_1$ is
in state $M$ and $c_2$ in state $I$ for address $a$, and the corresponding
directory states in L2 match with their actual states. $c_2$ gets a store
request from its processor and sends an upgrade request to L2. L2 sends a
downgrade to $S$ request to $c_1$. $c_1$ receives the downgrade to $S$ request
and downgrades to $S$, sending a downgrade response to L2. Let's say $c_1$
decides to evict address $a$ to make room for another address. $c_1$ sends a
downgrade to $I$ response to L2. Let's say the second response (downgrade to
$I$) reaches L2 first, violating Local-Property \ref{cRespFifo}. L2 sends a
upgrade to $S$ response to $c_2$. Let's say $c_2$ received the upgrade response
and changed state to $S$.  L2 then receives the second downgrade to $S$ response
from $c_1$.  Now L2 will change its directory state to $S$ for $c_1$ because it
received a downgrade to $S$ response last. Now let's say $c_2$ sent an upgrade
to $M$ request to L2 because of a store from its core. L2 will send an downgrade
to $I$ request to $c_1$, but $c_1$ drops the request since it's already in state
$I$.  Hence L2 will never receive a response from $c_1$ leading to a deadlock.
This is illustrated in Figure \ref{complicated}.

Let's illustrate the need for Local-Property \ref{reqNoOvertakeResp}. Let's say
$c_1$ is in state $S$ and $c_2$ in state $I$ for address $a$. Suppose $c_1$
receives a store request and sends an upgrade to $M$ request to L2. L2 sends
back an upgrade to $M$ response to $c_1$. $c_2$ gets a load request for address
$a$ and sends an upgrade to $S$ request to L2. L2 sends a request to $S$ to
$c_1$. Suppose Local-Property \ref{reqNoOvertakeResp} is violated and the
request to $S$ reaches $c_1$ before the response to $M$. $c_1$ will drop the
request since its already in state $S$, and L2 will never get a response leading
to a deadlock. This is illustrated in Figure \ref{lessComplicated}.

\begin{figure}\small
\begin{inv}
A response \Resp{p}{c}{a}{x}, where $p = c.parent$, can be sent from $p$ only
if
\begin{enumerate}
\item a thread that handling a request from $c$ to $p$ for
address $a$ is executing, and
\item $x > y$, where $p.dir[c][a] = y$ just before sending the response
\end{enumerate}
\label{cSendRespPre}
\end{inv}
\begin{inv}
A response \Resp{c}{p}{a}{x}, where $p = c.parent$, can be sent from $c$ only
if $x < y$, where $c.state[a] = y$ just before sending the
response\label{pSendRespPre1}
\end{inv}
\begin{inv}
If there is a thread in node $c$ that is waiting to receive a response from its
parent $p$ for an address $a$, then response \Resp{c}{p}{a}{x} can be sent only
on executing a thread that is handling a request from $p$ to $c$ for address
$a$. For instance, $a$ can not be evicted from $c$, if there is a thread
handling a request from $p$ to $c$ for address $a$.
\label{pSendRespPre2}
\end{inv}
\caption{Local-Properties for sending responses}
\label{sendResp}
\end{figure}

We will illustrate why Local-Properties \ref{cSendRespPre} and
\ref{pSendRespPre2} are important. Let's say $c_1$ is in $S$ and $c_2$ is in $I$
for address $a$ the corresponding directory states match with the actual states.

In the first scenario, let $c_1$ receive a store request for address $a$. It
sends an upgrade to $M$ request to L2. Then $c_1$ gets another load request from
the core for a different address $a'$ which is not present in $c_1$. If $c_1$
violates \ref{cSendRespPre} and evicts address $a$ to allocate a line for
address $a$, then it sends a downgrade to $I$ response for address $a$ to L2.
Meanwhile L2 has sent an upgrade to $M$ response to $c_1$. After sending the
response, L2 receives the downgrade to $I$ response form $c_1$. So L2 now
assumes that $c_1$ is in $I$ while in reality, $c_1$ is in $M$. This breaks
Invariant \ref{conservative} which creates the problems described earlier. This
is shown in Figure \ref{checkit}

In the second scenario, let's say L2 violates \ref{pSendRespPre2} and
voluntarily sends an upgrade to $M$ response for address $a$ to $c_1$.
Meanwhile, $c_1$ receives a load request from the core for a different address
$a'$ which is not present in $c_1$, and hence evicts $a$ to allocate a line for
$a$. $c_1$ then gets the upgrade to $M$ response from L2 while L2 gets the
downgrade to $I$ response, breaking Invariant \ref{conservative} again. This is
shown in Figure \ref{checkit2}.

In general, any crossing of response messages between a cache and its parent,
like the ones shown in Figures \ref{checkit} and \ref{checkit2} results in
violation of Invariant \ref{conservative}.

\begin{figure}\small
\begin{inv}
If a response \Resp{p}{c}{a}{x} is sent from $p$, then $p.dir[c][a] \gets x$\label{cSendRespPost}
\end{inv}
\begin{inv}
If \Resp{c}{p}{a}{x} is sent $c.state[a] \gets x$\label{pSendRespPost}
\end{inv}
\begin{inv}
On receiving a response \Resp{c}{p}{a}{x}, where $p = c.parent$, $p.dir[c][a]
\gets x$\label{pRecvResp}
\end{inv}
\begin{inv}
On receiving a response \Resp{p}{c}{a}{x}, where $p = c.parent$, $c.state[c]
\gets x$\label{cRecvResp}
\end{inv}
\begin{inv}
$c.state[a]$ can change only on $c$ sending a response to or receiving a
response from $c.parent$\label{cState}
\end{inv}
\begin{inv}
$p.dir[c][a]$ can change only on $p$ sending or receiving a response from
its child $c$\label{pState}
\end{inv}
\caption{Local-Properties for state changes}
\label{stateChange}
\end{figure}

\begin{figure}\small
\begin{inv}
\Req{p}{c}{a}{x}, where $p = c.parent$, can be sent only if $x < p.dir[c][a]$\label{pSendReqPre}
\end{inv}
\begin{inv}
If \Req{p}{c}{a}{x}, where $p = c.parent$, is received, if $x \ge c.state[a]$,
then the request is dropped\label{pSendReqPost}
\end{inv}
\begin{inv}
\Req{c}{p}{a}{x}, where $p = c.parent$, can be sent only if $x > c.state[a]$\label{cSendReqPre}
\end{inv}
\begin{inv}
If \Req{c}{p}{a}{x}, where $p = c.parent$, is received, it $x \le p.state[a]$,
then the request is dropped\label{cSendReqPost}
\end{inv}
\caption{Local-Properties for sending and receiving a request}
\label{sendReq}
\end{figure}

Local-Properties \ref{pSendReqPre} and \ref{pSendReqPost} are needed together
because an incoming request is dropped if the state is already lower than or
equal to the requested state without sending a response. So, the cache waiting
for a response will never receive one, causing a deadlock.  Similarly,
Local-Properties \ref{cSendReqPre} and \ref{cSendReqPost} are also needed
together.

\begin{figure}\small
\begin{inv}
An incoming request message from a child or the core should not block an
incoming request message from the parent, even for a different
address\label{cReqNoBlockPReq}
\end{inv}
\begin{inv}
An incoming request messages for an address $a$ should not block an incoming
response message for address $a$, even from a different
source.\label{reqNoBlockResp}
\end{inv}
\caption{Blocking restrictions}
\label{blocking}
\end{figure}

\begin{figure}\small
\begin{inv}
$p$ can not receive a request \Req{c}{p}{a}{x} from $c$, where $p = c.parent$,
if either
\begin{enumerate}
\item address $a$ has been chosen for eviction by some thread and that thread
has not yet replaced $a$, or
\item a thread in $p$ is handling a request for address
$a$ from any source
\end{enumerate}
\label{pHandleReq}
\end{inv}
\begin{inv}
$c$ can not receive a request \Req{p}{c}{a}{x} from $p$, where $p = c.parent$,
if either
\begin{enumerate}
\item address $a$ has been chosen for eviction by some thread and that thread
has not yet replaced $a$, or
\item a thread in $p$ is handling a request for address
$a$ from $p$, or
\item a thread in $p$ is handling a request for address
$a$ from one of $p$'s children $c'$, and that thread will not send any
(downgrade) request to any child for address $a$ anymore
\end{enumerate}
\label{cHandleReq}
\end{inv}
\caption{Local-Properties for receiving a request (used by the scheduler)}\label{recvReq}
\end{figure}

\begin{figure}\small
\begin{inv}
Address $a$ present in node $c$ can not be chosen for eviction by a thread
executing in $c$ if either
\begin{enumerate}
\item address $a$ has been chosen for eviction by some thread and that thread
has not yet replaced $a$, or
\item a thread in $c$ is handling a request for address
$a$ from any source
\end{enumerate}\label{evict}
\caption{Local-Properties for suspending a thread during choosing an address for eviction}
\end{inv}
\end{figure}

\begin{figure}\small
\begin{inv}
Each node has at least one dedicated thread resource for handling only messages
(requests or responses) from the parent\label{dedicate1}
\end{inv}
\begin{inv}
Each node has at least one dedicated thread resource for handling only responses
(from children or from the parent)\label{dedicate2}
\end{inv}
\begin{inv}
Each node has at least three thread resources\label{dedicate3}
\end{inv}
\caption{Dedicated resources in a node}
\end{figure}

Local-Properties \ref{evict}, \ref{pHandleReq} and \ref{cHandleReq} are needed
because once a thread has finished up-to a point in its execution, it can not
re-execute that portion. For instance, consider a thread handling an upgrade
request from a child. Let's say thread $t$ has progressed to the point where it
is simply waiting to send an upgrade response to the child. Now, if the line
gets evicted because it was chosen for replacement, or it got downgraded by the
parent, then thread $t$ can not re-execute its procedure again. To prevent this,
only one thread is allowed to operate on a particular address, except in the
case of downgrade requests from the parent. In this case, a thread is created to
handle the downgrade request from the parent even if there is another thread
handling a request for the same address from a child, if the latter thread is
just waiting for a response from the parent. This is done to avoid the deadlock
shown in the following scenario.

Let's illustrate the need for Local-Properties \ref{cReqNoBlockPReq},
\ref{dedicate1}, \ref{cReqNoBlockPReq} and \ref{dedicate2}. Let's say both $c_1$
and $c_2$ are in state $S$ for address $a$, and the corresponding directory
states in L2 match with their actual states. Both $c_1$ and $c_2$ get store
requests from the processor for address $a$, and they send upgrade to $M$
requests to L2. Let's say that L2 received the request from $c_2$ first.  L2
sends a downgrade to $I$ request to $c1$.

Let's first say Local-Property \ref{cReqNoBlockPReq} is not held, and instead some
request from the core is blocking $c_1$ from receiving the request. The $c_1$
will never respond back to L2. But L2 will never start handling the request from
$c_1$ because of Local-Property \ref{cHandleReq}. This will create a deadlock.
The same deadlock is also created when $c_1$ is out of thread resources to
create a thread to handle the downgrade request from $L2$ (violating Local-Property
\ref{dedicate1}).%Figure \ref{deadlockByPBlock} depicts this scenario.

Let's extend this scenario further. Instead of violating Local-Properties
\ref{cReqNoBlockPReq} or \ref{dedicate1} and $c_1$ not responding, let $c_1$
send a downgrade to $I$ response to L2. Let's instead assume that Local-Property
\ref{reqNoBlockResp} is violated in L2. So, the downgrade to $I$ response from
$c_1$ is not received by L2, leading to the same deadlock. In the same
manner, if Local-Property \ref{dedicate2} is violated, the response can not be
handled by L2 creating the same deadlock yet again.

The need for Local-Property \ref{dedicate3} can be shown easily. Let's say that a
cache has exactly 2 thread resources, one for handling messages from parents,
and one for handling responses from either the parent or the children. Now, if
it receives a request from one of its children, it can not use either of the
thread resources. Thus, at least one more thread resource is needed for
handling requests from children. This thread resource can also be used for
handling any type of incoming message.

\begin{figure}\small
\begin{inv}
Starvation-freedom: A thread that keeps becoming ready to execute will
eventually be scheduled to execute.\label{starvation}
\end{inv}
\end{figure}

Local starvation-freedom is required in each cache node. If a node never
executes a particular suspended thread, even if it keeps getting ready to
execute, then the thread will not progress, leading to a deadlock.