### 2012-11-21 paper/isca40format/DistributedMsi.tex
before arvind is going to re-org

\section{Distributed MSI}
\label{sec:DistributedMsi}

In order to transfer the Global MSI protocol into a distribute MSI protocol, we
need to make sure that a method can only read and write the state of the cache
with which the method is associated, and it affects the state of other caches
only by sending messages. In order to read the state of a child's cache we
maintain a shadow state of in the form of a directory. Thus the coherence state
of a cache $c$, in addition to the $c.state[a]$ field, also contains a
directory field containing the shadow state of each child's state field, \ie,
$c.dir[c'][a]$, for child $c'$.

We extend our sequential language for writing methods with two communication
primitives: \send{} $(Req|Resp|Data)\langle c_1 \rightarrow c_2, a,
(x|data)\rangle$ and \receive{} $(Req|Resp|Data)(\langle c_1 \rightarrow c_2,
a, (x|data)\rangle$ where caches $c_1$ and $c_2$ must satisfy the parent-child
relation. Both of these primitives are \emph{blocking}, that is, \send{} may
block because the network is clogged, \ie, has no buffer space to hold a
message and a \receive{} may block because the message has not arrived or cache
has no resources to process it. These primitives make our sequential language
methods ``suspensive'. Request or response messages have very different
characteristics and a data message is conceptually always associated with some
response message. We think of each \textsc{coreReq} as launching a thread in
L1. This thread in turn may get processed, send request or response messages,
may get suspended and eventually die. Thus potentially every request message
can launch a new thread in the cache that receives it. A response message,
however, can only die or generate more response messages; it can never generate
a request message. However a response message can also block waiting for the
corresponding data message. Each thread has an address associated with it which
corresponds to the message that created it.

In order to write the distributed MSI protocol using suspensive threads, we will
maintain the following sets of invariants. In the next section, we will show in
the requirements in the network and the bounds on the number of virtual channels
to avoid head of line blocking. We will also show the other resource
requirements, as well as the properties of the thread scheduler that will ensure
these invariants there.

\subsection{Relation between directory and state of the children's caches}
\begin{theorem}
Conservative: For two nodes $p, c$, where $p = c.parent$, $\forall a,
p.dir[c][a] \ge c.state[a]$.  \label{conservative}
\end{theorem}

This ensures that if a cache makes decision about compatibility of states of
its children using the local directory information, then it does not violate the
compatibility of the children's real states.

\subsection{Network properties}
Regarding the \send{} and \receive{} operations, \ie the network, the following
properties can assume the following:
\begin{theorem}
It a thread has to send a message, it will eventually be able to send the
message.
\label{canSend}
\end{theorem}

\begin{theorem}
If a cache sends an upgrade-to-$x$ request to its parent, it will eventually
get an upgrade-to-$y$ response from the parent where $y \ge x$.
\label{willRecv}
\end{theorem}

\begin{theorem}
If a cache sends an downgrade-to-$x$ request to one of its children, it will
eventually get a downgrade-to-$y$ response from the child where $y \le x$.
\label{willRecv2}
\end{theorem}

\subsection{Eviction}
If handling of a request requires eviction of another cache line, then too the
thread handling the request may get suspended. This would happen if all the
lines that can be allocated for this request may accessed by other suspended
threads. But we still assume the following invariant.

\begin{theorem}
If a thread is waiting for a line to be allocated, \ie for any address to become
available for eviction, then some address will eventually become available for
eviction
\label{evictDead}
\end{theorem}

It this is not true, a thread that is waiting for being allocated a free line
will never proceed leading to a deadlock.

\subsection{Data transfer}
Regarding the responses received, the designer can assume the following 2
invariants. These invariants are used to wait for receiving data messages
appropriately, as only some state transfers also cause data transfers.
We will show the exact need for these invariants later.

\begin{theorem}
If $p$ sends \Resp{p}{c}{a}{x} to $c$, where $p = c.parent$, then $p.dir[c][a]$
just before $p$ sends the response is the same as $c.state[a]$ just before $c$
receives the response.
\label{pcSame}
\end{theorem}

\begin{theorem}
If $c$ sends \Resp{c}{p}{a}{x} to $p$, where $p = c.parent$, then $c.state[a]$
just before $c$ sends the response is the same as $p.dir[c][a]$ just before $p$
receives the response.
\label{cpSame}
\end{theorem}

\subsection{Need for invariants}
The need for these network invariants is intuitive. If, for instance, a thread wants to
send a message in the output channel of the node and is not able to send one
forever, then the thread can not proceed further. If a thread is waiting for a
response (for a request previously sent), and it never gets the response, the
thread can never proceed further. These scenarios create a deadlock.

We will illustrate the need for Invariants \ref{pcSame}, \ref{cpSame} and
\ref{conservative} using an example of a 2-level system containing two L1
caches $c_1$ and $c_2$ with a shared L2 cache $p$.

Invariant \ref{pcSame} and \ref{cpSame} are important to make sure that the
data transfers happen correctly. Let $p.dir[c_1][a] = S$, $p.dir[c_2][a] = I$,
$c_1.state[a] = S$, and $c_2.state[a] = I$. $c_2$ gets a store request, and
sends an upgrade-to-$M$ to $p$. $p$ receives the request, but assumes that
$c_1$ already has data for $a$ since $p.dir[c_1][a] = S$. Invariant
\ref{pcSame} is violated. So, it does not transfer the data to $c_1$, resulting
in $c_1$ waiting for data, and hence deadlock. A dual condition can be
constructed similarly, in which $c_1$ downgrades its state for address $a$ from
$S$ to $I$, and the parent assumes that $c_1$ is in state $M$, and keeps
waiting for data.

Invariant \ref{conservative} is important because a parent node consults its
directory to decide the children it needs to send a downgrade request to.  Let
$p.dir[c_1][a] = S$, $c_1.state[a] = S$, $p.dir[c_2][a] = I$ but
$c_2.state[a] = S$. If $c_1$ gets a store request from the processor and sends
an upgrade-to-$M$ request to $p$, $p$ will assume that $c_2$'s state for address $a$
is also $I$ and hence send an upgrade-to-$M$ response to $c_1$. But since the real
state of  $c_2$ for address $a$ is $S$, it will break Single-writer Invariant
\ref{singleWriter}.

Thus, if all the invariants above (\ref{conservative}, to \ref{cpSame}) hold,
then the threads given in Figure \ref{realistic} can not deadlock because it
will always be able to a) send a request, b) send back a response, c) get back
a response for a request it sent, d) receive data when required. Though the
threads executing in a cache are actually suspensive, these invariants give an
illusion of the threads being non-suspensive barring the following exceptions.
Here $this$ represents the cache under under consideration.

\begin{enumerate}
\item $this.dir[c][a]$ can keep decreasing while a thread corresponding to
address $a$ is executing.
\item $this.dir[c][a']$ can keep decreasing while a thread which is evicting
line $a'$ is executing.
\item If a thread receives responses from the parent, then $this.state[a]$ can
decrease between the operation of sending a request to the parent and the
operation of receiving a response from the parent.
\end{enumerate}

As can be seen, these interactions are minimal, and can be ignored since they
affect neither the handling of upgrade nor downgrade requests. Hence, the
correctness of the protocol can be argued in terms of a single threaded,
non-suspensive, execution model. The formal proof for this, as well as the
invariants stated below is beyond the scope of this paper.

In order to design the distributed MSI protocol, the following considerations
must be kept in mind.
\begin{enumerate}
\item There are 3 kinds of incoming messages to be handled: request from that parent,
requests from children and responses from children.
\item A downgrade-to-$x$ message might be received from the parent even if the
state of the cache is $ \le x$. This has to be handled.
\item A parent sends a response to the child only when a request is received. But a child can send an voluntary/unsolicited response to the parent during evictions.
\item A request should be sent to the parent only if the current cache's state
is lower than what's desired to handle its incoming request. Similarly, a
request should sent to its child only if the directory's notion of that cache's
state is higher than what's desired.
\item A second request is not sent for the same address from the same cache to the same
destination, until a response is received for the first request.
\item A cache can not evict a line if it is currently waiting for a response from the parent for the address corresponding to the line.
\end{enumerate}

%We always reserve resources, \eg, cache line slot for an address before sending
%an upgrade request for that address.

The handling of messages follow the exact same algorithm as given for the
global MSI protocol in Figure \ref{msi-template} with the caveat that instead
of reading a child's state, the cache $this$ can only read its own directory.
So, it suffices to simply replace $c.state[a]$ by $this.dir[c][a]$ every time
the state of child $c$ is read by $this$.

As we mentioned above, a new handler has to be written for handling voluntary
responses. This is shown in Figure \ref{msi-unsolicited}.

\begin{figure}
\small
\begin{algorithmic}
\Proc {\dRespL}{$c , a, x$}
  \If {$isModified(this.dir[c][a])$}
    \State \receive{} \Data{c}{this}{a}{d};
    \State $this.data[a] \gets d$;
  \EndIf
\EndProc
\end{algorithmic}
\caption{Handling an voluntary response message from a child $c$, sent if the
child evicts line $a$ and downgrades the line to $x$}
\label{msi-unsolicited}
\end{figure}

\begin{figure}
\small

\begin{algorithmic}
\State // Upgrading $this$ by sending request to parent and getting a response
\Proc {\uReqL}{$p, a, x$}
  \If {$this$ is LLC \&\& $this.state[a] = I$}
    \State \send{} \Req{this}{Memory}{a}{M};
    \State \receive{} \Data{Memory}{this}{a}{d};
    \State $this.data[a] \gets d$;
    \State $this.state[a] \gets M$;
  \ElsIf
  \State \send{} \Req{this}{p}{a}{x};
  \State \receive{} \Resp{p}{this}{a}{z};
  \If {$this.state[a] = I$}
    \State \receive{} \Data{p}{this}{a}{d};
    \State $this.data[a] \gets d$;
  \EndIf
  \State $this.state[a] \gets z$;
  \EndIf
\EndProc
\State // Downgrading $this$' child by sending request to child and getting a
response
\Proc {\dReqL}{$c, a, x$}
  \State \send{} \Req{this}{c}{a}{x};
  \While {$this.dir[c][a] > x$}
  \State \receive{} \Resp{c}{this}{a}{z};
  \If {$isModified(this.dir[c][a])$}
    \State \receive{} \Data{c}{this}{a}{d};
    \State $this.data[a] \gets d$;
  \EndIf
  \State $this.dir[c][a] \gets z$;
  \EndWhile
\EndProc
\State // $this$ finally upgrading the directory state of its child and sending
a response to its child
\Proc {\uResp}{$c, a, x$}
  \State \textbf{if} ($this.dir[c][a] = I$)
  \State \;\;\;\; \send{} \Data{this}{c}{a}{this.data[a]};
  \State $this.dir[c][a] \gets x$;
\EndProc
\State // $this$ finally downgrading its state and sending a response to its
parent
\Proc {\dResp}{$p, a, x$}
  \State \textbf{if} ($isModified(this.state[a])$)
  \State \;\;\;\; \send{} \Data{this}{p}{a}{this.data[a]};
  \State $this.state[a] \gets x$;
\EndProc
\end{algorithmic}
\caption{Distributed MSI protocol. In all the methods $p$ is the parent of
$this$, $c$ is a child of $this$, and the methods work on address $a$ to change
its state (or directory) to $x$. This function only shows the auxiliary
functions and the handler for voluntary responses from children. The handler
for requests are exactly the same as in Figure \ref{msi-template}, with every
occurence of $c'.state[a]$ replaced by $this.dir[c'][a]$}
\label{realistic}
\end{figure}