### 2012-11-21 paper/isca40format/Invariants.tex

\section{Local properties that ensure global invariants}
\label{sec:properties}

In this section, we will show some of the core properties that are extracted
from the requirements spelt out in Section \ref{sec:network} as well as the
implementation of the distributed protocol's auxiliary functions given in Figure
\ref{realistic}. These properties can be independently used to prove the
invariants presented in Section \ref{sec:DistributedMsi} (Invariants
\ref{conservative} to \ref{cpSame}). We do not present the proof here, but give
some intuition. Figure \ref{sendReq} lists these local properties.

\floatstyle{boxed}
\restylefloat{figure}

\begin{figure}\small
\begin{inv}
A response \Resp{p}{c}{a}{x}, where $p = c.parent$, can be sent from $p$ only
if
\begin{enumerate}
\item a request \Req{c}{p}{a}{z} was received from $c$,
\item $p$ is not waiting for any response for address $a$ from $c$, and
\item $x > y$, where $p.dir[c][a] = y$ just before sending the response
\end{enumerate}
\label{pSendRespPre}
\end{inv}
\begin{inv}
A response \Resp{c}{p}{a}{x}, where $p = c.parent$, can be sent from $c$ only
if $x < y$, where $c.state[a] = y$ just before sending the
response\label{cSendRespPre1}
\end{inv}
\begin{inv}
If $c$ has sent a request \Req{c}{p}{a}{x}, then $c$ can send a response
\Resp{c}{p}{a}{y} only on receiving a request from $p$ for address $a$.
\label{cSendRespPre2}
\end{inv}
%\caption{Local-Properties for sending responses}
%\label{sendResp}
%\end{figure}
%
%\begin{figure}\small
\begin{inv}
If a response \Resp{p}{c}{a}{x} is sent from $p$, then $p.dir[c][a] \gets x$\label{cSendRespPost}
\end{inv}
\begin{inv}
If \Resp{c}{p}{a}{x} is sent $c.state[a] \gets x$\label{pSendRespPost}
\end{inv}
\begin{inv}
On receiving a response \Resp{c}{p}{a}{x}, where $p = c.parent$, $p.dir[c][a]
\gets x$\label{pRecvResp}
\end{inv}
\begin{inv}
On receiving a response \Resp{p}{c}{a}{x}, where $p = c.parent$, $c.state[c]
\gets x$\label{cRecvResp}
\end{inv}
\begin{inv}
$c.state[a]$ can change only on $c$ sending a response to or receiving a
response from $c.parent$\label{cState}
\end{inv}
\begin{inv}
$p.dir[c][a]$ can change only on $p$ sending or receiving a response from
its child $c$\label{pState}
\end{inv}
%\caption{Local-Properties for state changes}
%\label{stateChange}
%\end{figure}
%
%\begin{figure}\small
\begin{inv}
\Req{p}{c}{a}{x}, where $p = c.parent$, can be sent only if $x < p.dir[c][a]$\label{pSendReqPre}
\end{inv}
\begin{inv}
If \Req{p}{c}{a}{x}, where $p = c.parent$, is received, if $x \ge c.state[a]$,
then the request is dropped\label{pSendReqPost}
\end{inv}
\begin{inv}
\Req{c}{p}{a}{x}, where $p = c.parent$, can be sent only if $x > c.state[a]$\label{cSendReqPre}
\end{inv}
\begin{inv}
If \Req{c}{p}{a}{x}, where $p = c.parent$, is received, it $x \le p.state[a]$,
then the request is dropped\label{cSendReqPost}
\end{inv}
\begin{inv}
A request can be sent by a cache node $n$ for address $n$ to another cache node $m$ only if
no $n$ is not waiting for a response for address $a$ from $m$.
\label{nodoublereq}
\end{inv}
\caption{Local-Properties useful in proving the correctness of MSI protocol}
\label{sendReq}
\end{figure}

 We will use the violation of Local-Properties \ref{pSendRespPre} and
\ref{cSendRespPre2} as examples to illustrate the importance of local
properties.  In our usual system, let's say $c_1.state[a] = p.dir[c_1][a] = S$,
$c_2.state[a] = p.dir[c_2][a] = I$.

\floatstyle{plain}
\restylefloat{figure}
\begin{figure}
\centering
\includegraphics[scale=0.4]{checkit}
\caption{Effect of violating Local-Property \ref{cSendRespPre2} (child evicts a ``pending'' line)}
\label{checkit}
\end{figure}
\floatstyle{boxed}
\restylefloat{figure}

Let $c_1$ receive a store request for address $a$. It sends an upgrade-to-$M$
request to $p$. Then $c_1$ violates \ref{cSendRespPre2} and evicts $a$ to
replace it with another address. It sends a downgrade to $I$ response for
address $a$ to $p$.  Meanwhile $p$ has sent an upgrade-to-$M$ response to
$c_1$. $c_1$ receives the response and upgrades address $a$ to $M$. $p$ then
receives the downgrade to $I$ response form $c_1$. So $p$ now assumes that
$c_1$ is in $I$ while in reality, $c_1$ is in $M$. This breaks Invariant
\ref{conservative} which creates the problems described earlier. This is shown
in Figure \ref{checkit}

\floatstyle{plain}
\restylefloat{figure}
\begin{figure}
\centering
\includegraphics[scale=0.4]{checkit2}
\caption{Effect of violating Local-Property \ref{pSendRespPre} (parent sends a voluntary response)}
\label{checkit2}
\end{figure}
\floatstyle{boxed}
\restylefloat{figure}

Let's say $p$ violates \ref{pSendRespPre} and voluntarily sends an
upgrade-to-$M$ response for address $a$ to $c_1$.  Meanwhile, $c_1$ evicts $a$
to replace it with another address. $c_1$ then gets the upgrade-to-$M$ response
from $p$ while $p$ gets the downgrade to $I$ response, breaking Invariant
\ref{conservative} again. This is shown in Figure \ref{checkit2}.

In general, any if response messages from both a cache to its parent and from
the parent to the same cache are in flight together, like the scenarios in
Figures \ref{checkit} and \ref{checkit2} Invariant \ref{conservative} will be
violated.

\begin{figure}
\begin{subfigure}{.25\linewidth}
\centering
\begin{tabular}{|ccc|}
\hline
$M$ & $\rightarrow$ & $I$\\
$O$ & $\rightarrow$ & $S$\\
$S$ & $\rightarrow$ & $O$\\
$I$ & $\rightarrow$ & $I$\\
\hline
\end{tabular}
\subcaption*{$toCompatible$}
\end{subfigure}
\begin{subfigure}{.25\linewidth}
\centering
\begin{tabular}{|c|c|}
\hline
$M$ & $\checkmark$\\
$O$ & $\checkmark$\\
$S$ & $\times$\\
$I$ & $\times$\\
\hline
\end{tabular}
\subcaption*{$dataToUpper$}
\end{subfigure}
\begin{subfigure}{.48\linewidth}
\centering
\begin{tabular}{|c|cccc|}
\hline
& $M$ & $O$ & $S$ & $I$\\
\hline
$M$ & $=$ & $>$ & $>$ & $>$\\
$O$ & $<$ & $=$ & $>$ & $>$\\
$S$ & $<$ & $<$ & $=$ & $>$\\
$I$ & $<$ & $<$ & $<$ & $=$\\
\hline
\end{tabular}
\subcaption*{$<$ (and other) relations}
\end{subfigure}
\caption{MOSI protocol mappings and $<$ relations}
\label{mosi}
\end{figure}

These properties do not require messages to physically come from the sources as
indicated by the requests or responses -- some other cache node can forward
these messages. These can be used to generalize our procedure to
cache-intervention based protocols like MOSI. We generalize the states of MOSI
using the $<$ relation and auxiliary functions $maxCompatible$ and $isModified$,
as shown in Figure \ref{mosi}.

In a hierarchical setting for the MOSI protocol, any cache can send data
with the response granting permissions to the requesting L1 cache, instead of
going through the hierarchy.  On a load request, L1 cache $c$ sends an
upgrade-to-$O$ request, instead of an upgrade-to-$S$ request. The request keeps
going up the hierarchy, sending an upgrade-to-$O$ request to each of the parent
on the way, till it reaches a cache $p$ which has that line in $M$ or $O$
state. If some other child is in $O$ or $M$ state, this node sends a
upgrade-$c$-to-$O$ in addition to a downgrade-to-$S$ request to that cache.
This upgrade-$c$-to-$O$ + downgrade-to-$S$ keeps propogating down the tree till
it reaches a node $n$ whose children are all in a state $< O$. $n$ then
downgrades its state, forwards the upgrade-to-$O$ response to $c$ and
simultaneously sends a fwd-and-downgrade-to-$S$ response to its parent, which
keeps forwarding that till it reaches $p$. If $c$ gets a store request, it will
send an upgrade-to-$M$ request, and $p$ would send a downgrade-to-$I$ request
instead, to its children.

The caches encountered on the way from $n$ to $p$ should handle an extra
scenario where it gets a downgrade response, without a forward response.In this
case, the data and permission forwarding has to be done from this cache to $c$.
It is the designer's responsibility to ensure that this case is handled.

This protocol obeys all the local properties and the requirements mentioned in
this section except that a request sent to a cache might overtake a response
forwarded to that cache. To avoid this, once the forwarding message is
delivered to the intended destination ($c$ in our case), then $c$ sends back an
ACK. $c$'s parent is allowed to to start sending downgrade requests for that
address only after receiving this ACK. This ACK is sent back to each of the
parents till it reaches $p$, which allows each of these caches to send
downgrade requests for that address towards $c$.  This satisfies all the
requirements of Section \ref{sec:network} and the local properties given in
Figure \ref{sendReq}, thereby automatically being correct. The reasoning for
its correctness gets localized to checking if the ordering Requirement
\ref{reqNoOvertakeResp} is maintained.