### 2012-11-22 paper/isca40format/Conclusion.tex
before reorging scheduler etc

\section{Conclusions}
\label{sec:conclusion}

We have presented an automatic procedure to generate a distributed cache
coherence protocol from a centralized protocol. We have shown how this can be
implemented realistically and give the lower bound on the number of virtual
channels to make this coherence protocol deadlock-free and the storage
requirements. We have also provided intuition for the correctness of the
generated protocol. We have shown that our method of designing protocols
substantially reduces the required effort in implementing provably correct
distributed protocols.