### 2014-01-07 paper/Title.tex

%\begin{titlepage}

\title{Stuff
}

%       {\Large (Regular Submission)}}

%\author{Names withheld for submission}%\vspace{30pt}}

\author{
%    Muralidaran Vijayaraghavan\\Massachusetts Institute of Technology\\vmurali@csail.mit.edu\and
%    Nirav Dave\\SRI International\\ndave@csl.sri.com\and
%    Adam Chlipala\\Massachusetts Institute of Technology\\adamc@csail.mit.edu\and
%    Arvind\\Massachusetts Institute of Technology\\arvind@csail.mit.edu
}

\date{}

\maketitle % \thispagestyle{empty}

%=========================================================================
%  Abstract
%=========================================================================

\begin{abstract}
\input{Abstract}
\end{abstract}

%\bigskip

%\centerline{\footnotesize {\bf Keywords}: Guarded Atomic Actions, HW Synthesis, Modular Composition}

%\end{titlepage}