### 2014-01-30 paper/ProofStructure.tex
System is a little more motivated. Pull out bits of Proof structure to end and drop into new section

\section{Understanding the Correctness of the Protocol}
\label{Sec:ProofStructure}

For our protocol to be correct it must conform to our notion of
correct behavior \eg{} store atomicity Intuitively, we can understand
the correctness at a high-level from the following two facts.  that
particular address. Second, that any child node upgrades its local
state from \In to either \Sh or \Mo, it will get the same version of
data as it's parent which is the current global value.

%%%%%%%%%%%%%%%%%

Now that we have established a high-level view of the base protocol,
we will concretely discuss the structure of the proof. Much of the
structure discussed is necessary for the proof of correctness only and
not the actual implementation, \eg{} time. These additional components
can easily be optimized away as they're values are not used in the
actual handling of cache operations.