### 2014-02-03 paper/Abstract.tex
stuff

Correct implementation of cache coherent systems is of high importance in
industrial use. Model-checking is inadequate for verifying a large system with
hierarchical caches because of its size and complexity. We provide a
theorem-prover verified proof for the safety properties of a size-parameterized
hierarchical MSI-based coherent system. The system is specified in terms of
atomic transitions accessing only local state.  Previously, it has been shown
that these atomic transitions can be translated automatically into efficient
hardware. Verifying the specification based on atomic transitions therefore
guarantees the correctness of the actual implementation.
The proof is built in a modular fashion, identifying key invariants common with
other protocols like MESI, MOSI, \etc. We informally show how these other
protocols can be proved reusing modules from the proof of the MSI protocol.