### 2014-02-04 paper/Background.tex

\section{Correctness of a Cache Coherence Protocol}
\label{Background}

Before we can discuss our protocol and proof framework we must first establish
exactly what it means for a cache coherence protocol to be correct.

Intuitively, the simplest view of shared memory is as a single ported memory
without caches multiplexed amongst the many individual clients/processors. In
this view, all operations in the system appear to have a single unique total
ordering.

All modern systems have caches between the processor and memory for faster
memory access.  It is often also useful to permit some reorderings of requests
from a processor to the memory to improve performance.  For instance, in the
context of a processor with local store buffer, we may wish to allow a store to
bypass a subsequent load request to a different location. The specification of
which reorderings are permitted or restricted defines the memory consistency
model.

Most memory consistency models can be partitioned into two
parts~\cite{Arvind-memory-model}: a) a ``global'' store-atomicity property
stating that for each location/address in memory, every read and write access
to that location occurs in some total order with each operation seeing the
effects of previous operations and not later ones (in other words, each
location behaves like an atomic register) and b), a ``local'' reordering
restriction between requests to different locations within each processor.
Changing the memory model by modifying the reordering restrictions will not
incur a reverification of the store-atomicity property.

The mechanism employed by the memory sub-system to ensure store-atomicity per
location, in the presence of caches, is known as \emph{cache coherence}. In the
presence of caches, multiple copies of the same address can exist in different
caches. If these copies are independently written, the same address can end up
having different values in different caches, violating the atomic register-like
behavior unless the caches co-ordinate via a \emph{cache coherence protocol}.

We will now provide the formal definition of store atomicity.

\subsection{Formal definition of store atomicity}
We will model an abstract memory sub-system, which interfaces with a set of
processors. The requests issued by the processors to the memory sub-system can
be modeled using a function \reqFn. $\reqFn(p, i)$ returns the $i^{th}$ request
issued by processor $p$.

%Figure~\ref{req-resp} shows the
%data-type corresponding to a request issued from the processor to the
%memory sub-system.

A request contains the address (\addrQ) corresponding to the request, and a
\desc{} indicating whether the request is a load or a store. A store request
also contains the data supplied by the processor for that address (\dataQ). A
response contains a label (\labelR) containing a processor name and a position
to uniquely identify the request.  In the case of a load request, the memory
sub-system returns \dataR, the data returned by the memory sub-system for that
address. The memory sub-system also returns an $\mathbb{N}$, \timeR, which
intuitively represents the position at which the request is handled among the
global set of handled requests from all the processors. \timeR{} is used solely
for the definition of store atomicity, a real system does not return such a
field.

%We model request of a processor formally with the function
%\reqFn. $\reqFn(p, i)$ returns the $i^{th}$ request issued by
%processor $p$, \ie{} $\reqFn: (\textit{Processor} * \mathbb{N}) @->
%\Request$.  The processor name and the position of the request within
%the processor uniquely identifies a request.

%\begin{figure}
%\centering
%\begin{subfigure}{5.5cm}
%\begin{tabular}{|lp{4.5cm}|}
%\hline
%\multicolumn{2}{|c|}{\Request}\\
%\hline
%\addrQ:& Address associated with the request\\
%\desc:& Request type (\Ld{} or \St)\\
%\dataQ:& Data associated with a \St{} request\\
%\hline
%\end{tabular}
%\end{subfigure}~~~~
%\begin{subfigure}{5.5cm}
%\begin{tabular}{|lp{4.5cm}|}
%\hline
%\multicolumn{2}{|c|}{\Response}\\
%\hline
%\labelR:& Label identifying the request corresponding to this response\\
%\timeR:& Natural number representing global order of (processing) requests\\
%\dataR:& Data returned by the memory sub-system for a \Ld{} request\\
%\hline
%\end{tabular}
%\end{subfigure}
%\caption{Data types for request and response}
%\label{req-resp}
%\end{figure}

The \Response{} set contains the set of responses from the memory sub-system.
Responses should obey the following sanity conditions. 
\begin{defn} (The labels in the responses are unique)
%Since \dataR{} field is
%irrelevant for store requests, so uniqueness of responses does not depend on
%\dataR{} field for stores.)
\small
\begin{multline*}
\textit{uniqRespLabels} := \forall (r_1, r_2 \in \Response),\\
r_1.\labelR = r_2.\labelR @-> r_1.\timeR = r_2.\timeR \wedge\\
(\reqFn(r_1.\labelR) = \Ld @-> r_1.\dataR = r_2.\dataR)
\end{multline*}
\end{defn}

\begin{defn} (No response can share a \timeR{} value with another store response.)
\small
\begin{multline*}
\textit{uniqStTimes} := 
\forall (r_1, r_2 \in \Response),
r_1.\timeR = r_2.\timeR @-> \\
(\reqFn (r_1.\labelR)).\addrQ = (\reqFn (r_2.\labelR)).\addrQ @->\\
(\reqFn (r_1.\labelR)).\desc = \St @->
r_1.\labelR = r_2.\labelR
\end{multline*}
\end{defn}

In addition, the global ordering of responses to the same processor should not
violate the local ordering of the corresponding requests.

\begin{defn}
\small
\begin{multline*}
\textit{localOrdering} :=
\forall (r_1, r_2 \in \Response), (r_1.\labelR).\fst = (r_2.\labelR).\fst @->\\
(r_1.\labelR).\snd < (r2.\labelR).\snd @-> r_1.\timeR \le r_2.\timeR
\end{multline*}
\end{defn}

This brings us to the main condition for store atomicity which states that the
data returned for a load response corresponds to the \emph{last earlier} store
request from any processor that has been handled for that address,
\emph{last} and \emph{earlier} being defined according to the field \timeR{}
returned in the responses.  This proposition is indicated by
\textit{lastEarlierValue}$(a, t, d)$ for a load request for address $a$ handled
at time $t$, returning data $d$.  In case no store has happened earlier for a
particular address when a load is processed, then the load response returns the
initial value for the address, the whole proposition indicated by
\textit{initialValue}$(a, t, d)$.
\begin{defn}
\small
\begin{multline*}
\textit{initialValue} := \lambda (a \in Addr)\; (t \in \mathbb{N}) \; (d \in Data)\Rightarrow
d = \initData (a) \wedge \\
\forall (r' \in \Response), 0 \le r'.\timeR < t @->\\
\neg ((\reqFn (r'.\labelR)).\addrQ = t \wedge (\reqFn (r'.\labelR)).\desc = \St)
\end{multline*}
\end{defn}

\begin{defn}
\small
\begin{multline*}
\textit{lastEarlierValue} := \lambda (a \in Addr)\; (t \in \mathbb{N}) \; (d \in Data)\Rightarrow
\exists (r_m \in \Response),\\ d = (\reqFn (r_m.\labelR)).\dataQ \wedge 
(\reqFn (r_m.\labelR)).\desc = \St \wedge\\ (\reqFn (r_m.\labelR)).\addrQ = a \wedge
\forall (r' \in \Response), r_m.\timeR \le r'.\timeR < t @->\\
\neg ((\reqFn (r'.\labelR)).\addrQ = a \wedge (\reqFn (r'.\labelR)).\desc = \St)
\end{multline*}
\end{defn}

Using the above definitions, we can define the store-atomicity theorem as follows:
\begin{thm}
\small
\begin{multline*}
storeAtomicity := 
\textit{uniqRespLabels} \wedge
\textit{uniqStTimes} \wedge \textit{localOrdering} \wedge \\
(\forall (r \in \Response), \mylet q := \reqFn(r.\labelR) \myin q.\desc = \Ld \wedge \\
(\textit{initialValue}(q.\addrQ, r.\timeR, r.\dataR) \vee \textit{lastEarlierValue}(q.\addrQ, r.\timeR, r.\dataR)))
\end{multline*}
\end{thm}