### 2014-02-04 paper/Conclusion.tex

\section{Discussion}

We illustrated a direct realization of our proposed proof framework
for a very flexible cache coherence protocol implementation codifying
the intuition behind the system's correctness as simple
invariants. While this was took significant expertise, once done, we
were able to show that many key hardware modifications,\eg\ caching
policies, could be made with zero or very small natural changes to the
proof. These bits were small enough that an informed hardware designer
implementing our protocols can reasonably be expected to provide such
information during refinement giving them a viable path for producing
a provably correct performant cache coherence engine implementation.

The protocol framework we showed was very flexible, and could easily
be extended for variants that go beyond microarchitectural variants,
\eg\ silent drops, extension to MOSI with a relatively small amount of
additional reasoning.  This gives us good confidence that we can
effectively produce frameworks for other classes protocol,
\eg\ update-based protocols, non-inclusive cache hierarchies. Further,
nothing about our methodology requires we limit ourselves to cache
coherence systems; this approach should naturally extend to any
hardware-based system.

While fundamentally this work meets our goal of carrying verification
into through the design refinement step, there are still some
important mechanization to make this ready for industrial
use. Foremost, the simple transliteration between the Coq transition
system and the Bluespec hardware description implementation should be
fully mechanized. Further, the proposed extension to provide fairness
of transaction scheduling in the Bluespec compiler should be
implemented. Finally, to fully guarantee correctness of the resulting
hardware FSM, we should also fully verify the hardware generation as
in the FeSi~\cite{Tomas} efforts.