### 2014-02-04 paper/Deadlock.tex

\section{Liveness properties}
\label{liveness}

In Section~\ref{safety} we showed how the MSI protocol of
Figure~\ref{trans} obeys the store atomicity theorem, showing that
only correct behavior occurs. To complete our correctness picture, we
must also show that we avoid deadlock, \ie\ we make forward progress.

Forward progress can be either of the following:
\begin{enumerate}
\item Weak forward progress: The system never reaches a state where
  none of the transitions of Figure~\ref{trans} apply or
\item Strong forward progress: At any time, eventually a new request
  from some processor will be processed
\end{enumerate}

To prove strong forward progress given weak fairness, we need some
guarantees on the choice of executions of the transitions
Figure~\ref{trans} in each cache. Specifically, caches must ensure
local starvation-freedom in the processing of messages when sufficient
resources are available. This can be enforced via during hardware
compilation; for this paper we will focus only on weak forward
progress.

In order to establish that some transition always executes, we prove the
following important invariant.

\begin{inv}
\textit{waitCond}: When a cache $x$ has sent a request to another cache $y$ and
is waiting for a response from $y$, then either $y$ has not yet received the
request, or it has received the request and sent the appropriate response to
$x$.
\end{inv}

This condition ensures that every request gets a response eventually. Using
this invariant, it is easy to see that there will always be some cache in the
system which can either send a request to another cache (ChildSendReq or
ParentSendReq) or process a request (ParentRecvReq, ChildRecvReq or
ChildDropReq) or accept a response from another cache (ChildRecvResp or
ParentRecvResp). We do not give the formal proof here due to lack of space.