### 2014-02-05 paper/Deadlock.tex
intuition

\section{Liveness properties}
\label{liveness}

In Section~\ref{safety} we showed how the MSI protocol of Figure~\ref{trans}
obeys the store atomicity theorem. To complete the picture, we must also show
that it avoids deadlock, \ie{} the protocol makes forward progress.

Forward progress can be either of the following:
\begin{enumerate}
\item Weak forward progress: The system never reaches a state where
  none of the transitions of Figure~\ref{trans} apply, or
\item Strong forward progress: At any time, eventually a new request
  from some processor will be processed.
\end{enumerate}

To prove strong forward progress, we need some guarantees on the choice of
executions of the transitions Figure~\ref{trans} in each cache. Each cache must
ensure that it does not avoid receiving a particular available request from one
of its children indefinitely, when the guard for the corresponding transition
(ParentRecvReq) is true. Besides, all the transitions in Figure \ref{trans} are
voluntary in nature. Thus, the system can execute a sequence of transitions
which simply moves, say, the \Mo{} permission among the children of a cache,
without doing any further work with that permission
%This can potentially be enforced via during hardware
%compilation;
For this paper we focus only on weak forward progress, as we have not modeled
the algorithm for correctly choosing the transitions to guarantee strong
forward progress in Coq.

In order to establish that some transition always executes, we prove the
following important invariant.

\begin{inv}
\textit{waitCond}: When a cache $x$ has sent a request to another cache $y$ and
is waiting for a response from $y$, then either $y$ has not yet received the
request, or it has received the request and sent the appropriate response to
$x$.
\end{inv}

This condition ensures that every request gets a response eventually. Using
this invariant, it is easy to see that there will always be some cache in the
system which can either send a request to another cache if the cache is not
waiting for any response from the latter cache (ChildSendReq or ParentSendReq),
or accept a response from another cache if such a response is available
(ChildRecvResp or ParentRecvResp). But if all the caches are waiting for a
response, and no response is available, then the leaf caches can execute
ChildRecvReq transition (as the leaf does not have any children, and this
transition is executed even if a cache is waiting for a response from its
parent). Therefore, some transition can always be executed, proving weak
forward progress.