### 2014-02-05 paper/Intuition.tex
A pass over Intuition

\section{Proof of Store Atomicity for the MSI Protocol}
\label{Sec:Safety}

In this section, we present the structure of the proof of Theorem
\fullref{storeAtomicity} for a system implementing the MSI protocol of
Figure~\ref{trans}. We reduce the whole system proof to that of a set
of invariants between any parent-child pair. These invariants in turn
depend on two smaller invariants that encapsulate the core aspects of
correctness of the MSI protocol and allow the large scope of verified
hardware refinements we present later (Section~\ref{Sec:Refinements}).

The invariants hold after any number of executions of the transactions
starting from the initial state. We write ``time $t$'' for
the global state reached after firing $t$ transitions.

While we have proofs in Coq for each invariant, our discussion will be mainly
informal for brevity. A full description can be found at~\xxx{MURALI PUT
REFERENCE HERE}.
\adamc{I presume we'll want to submit a tarball of the Coq source code with the paper.
Is that what you have in mind here?}

\subsection{Reducing Correctness to Node-Level Invariants}

Store atomicity is a whole-system property. Intuitively our
protocol is correct because it guarantees at most one cache may
write at a time, and that any cache permitted to read has
the ``latest'' value. We formalize this intuition with the following 4
invariants.

\begin{inv}
\textit{latestValue}:
At any time $t$, a cache $c$ that is clean for an address $a$ will have the most
up-to-date value for that address, \ie{}
\begin{enumerate}
\item $c.\data[a] = \initData(a)$ and no store request $q'$ has been processed at
any time $t'$ such that $t' < t \wedge a = q'.\addrQ$, or
\item There is a store request $q$ that was processed at time $t_q$ such that
$t_q < t \wedge a = q.\addrQ \wedge c.\data[a] = q.\dataQ$ and no other store
request $q'$ was processed at any time $t'$ such that $t_q < t' < t \wedge a =
q'.\addrQ$.
\end{enumerate}
\label{latestValue}
\end{inv}
A clean cache is defined as follows.
\begin{defn}
\textit{clean}: A cache is said to be \clean{} for an address $a$ if and only if
the state of the cache for that address is at least \Sh{} and the directory 
state for all its children is at most \Sh.
\end{defn}

\begin{inv}
\textit{nonAncestorCompatible}: For two distinct caches $c_1$ and $c_2$ such
that neither is an ancestor of the other in the cache tree
hierarchy (where the ancestor relation is the reflexive-transitive closure of the parent relation), for each
address $a$, $c_1.\state[a]$ and $c_2.\state[a]$ are \textit{compatible}.
\label{nonAncestorCompatible}
\end{inv}
Compatibility of states is defined as follows.
\begin{defn}
\textit{compatibile}: Values of states $x_1$ and $x_2$ are \textit{compatible} iff
whenever either $x_1$ or $x_2$ is \Mo{}, the other is \In. Note that any two states $y_1 \le x_1$ and $y_2 \le x_2$ are also compatible if $x_1, x_2$ are compatible.
\label{compatible}
\end{defn}

%\begin{inv}
%\textit{processReq}: Load and store requests can be processed at a leaf cache
%only if it is at least in the \Sh{} and \Mo{} states, respectively, for that
%address.
%\label{processReq}
%\end{inv}
%
%\begin{inv}
%\textit{noReqAgain}:
%\item For a leaf cache $c$, $c.\pos$ changes only on processing a request. In
%that case, $c.\pos$ gets incremented.
%\label{noReqAgain}
%\end{inv}

\adamc{Something needs to be fixed up here, for consistency in how many invariants there are.}

Given these invariants, we codify our intuition by showing that
Theorem~\fullref{storeAtomicity} follows naturally from the
invariants.  A store may happen only when a leaf cache is in the \Mo{} state
for that address (Invariant~\fullref{processReq}). Leaf caches cannot
be ancestors of any other node, so by Invariant
\fullref{nonAncestorCompatible}, if a leaf cache is in \Mo{} state, all
other leaf caches will be in \In{} state preventing the possibility of
load and store requests for the same address (by Invariant
\fullref{processReq}), showing Definition~\fullref{uniqStTimes} holds.

Given Invariant~\fullref{noReqAgain}, once a response has been produced
for a request $i$, the cache may subsequently only process a $j^{th}$
request such that $j > i$. Thus there is at most one response to each
request, satisfying Definition~\fullref{uniqRespLabels}. A load request
may be processed only if a cache is at least in \Sh{} state (Invariant
\fullref{processReq}). As leaf caches have no children, a leaf cache in
at least the \Sh{} state for address $a$ is \clean{} for that address. A
load response supplies the data for that address present in the cache,
which, by Invariant~\fullref{lastestValue}, contains the ``latest
value.'' Thus, the final condition for Theorem~\fullref{storeAtomicity}
is also satisfied. All that remains is to show that the system maintains the
invariants.

\subsection{Reducing to Pairwise Invariants}
\label{pairwise}

Correctness of Invariants \fullref{localCompatibility} and \fullref{latestValue} can be
proved using the following invariants on a parent-child pair.

\begin{inv}
\textit{conservative}: The state of an address in a cache is never greater than
the corresponding entry in its parent's directory.
\label{conservative}
\end{inv}

\begin{inv}
\textit{localCompatibility}:
\begin{enumerate}
\item The state of a cache is never less than its directory entry for any of
its children.
\item Let $c_1$ and $c_2$ be distinct caches with the same parent $p$. Then,
$p.\dstate[c_1][a]$ and $p.\dstate[c_2][a]$ are \textit{compatible}.
\end{enumerate}
\label{localCompatibility}
\end{inv}

\begin{inv}
\textit{noTransitWrite}: Whenever data for an address $a$ is in transit (\ie{}
$\forall t, t_s \le t \le t_r$ where $t_s$ is the time of sending the data and
$t_r$ the time of receiving the data), no cache can process a store request for
$a$.
\label{noTransitWrite}
\end{inv}

For any two caches $c$ and $p$, such that $p$ is an ancestor of $c$ in
the cache tree hierarchy, from Invariants \fullref{conservative} and
\fullref{localCompatibility}, we can show that for any address $a$,
$c.\state[a] \le p.\state[c][a]$. The reason is that the ancestor
relation is the reflective-transitive closure of the parent relation,
and Invariants \fullref{conservative} and the first part
of \fullref{localCompatibility} imply a $\ge$ relation between the
states of a parent and its child.

Any two caches $c_1, c_2$ neither of which are an ancestor of the
other will have a lowest common ancestor $x$ which is an ancestor of
both $c_1$ and $c_2$ and has two distinct children $x_1$ and $x_2$
such that they are the ancestors of $c_1$ and $c_2$ respectively. For
any address $a$, $x.\dstate[x_1][a]$ and $x.\dstate[x_1][a]$ are
compatible by the second part of
Invariant~\ref{localCompatibility}. From this, it straightaway follows
that $x_1.\state[a]$ and $x_2.\state[a]$ are compatible (from
Invariant~\ref{conservative}). From the previous paragraph, it follows
that $c_1$ and $c_2$ are also compatible. This proves
Invariant \fullref{nonAncestorCompatibility}.

We prove Invariant \fullref{latestValue} by strong induction on
time. Consider two cases: cache $c$ is already \clean{} for an address
$a$ at time $t$ or the contrary.  If \clean{}, then it has the latest
value at time $t$. Since none of its children's directory entries can
be in \Mo{}, all its descendants should be no greater than \Sh{} for
that address (as discussed above). Any non-ancestor of $c$ and
non-descendant of $c$ must be compatible (as discussed above). Since
$c$ is \clean, it is at least in the \Sh{} state, so compatibility
means that non-descendants cannot be in \Mo{} state. All of $c$'s
ancestors are non-leaves, so they cannot process any store
request. Therefore, no cache save $c$ may have processed a store
request, which means that $c$ contains the latest value for address
$a$ at time $t+1$ if it was \clean{} at time $t$.

If $c$ was not clean at time $t$, it must have received data from
either its parent or its child.  As can be seen in Figure~\ref{trans},
whenever data is sent either from a parent to a child or from a child
to the parent, the sender is \clean. By the induction hypothesis, it
means that the sender has the latest value for the address. Because of
Invariant~\ref{noTransitWrite}, no store request can be processed when
the data is in transit, which means $c$ will have the latest value for
address $a$ at time $t+1$.

\subsection{Proving Parent-Child Hierarchy Properties}

%ndave: This needs some repointing

\adamc{This section seems like a prime place to reduce detail to save space, while also saving readers from trying to understand proofs that are mechanized, anyway.}

We present two basic invariants that the MSI protocol of Figure~\ref{trans}
obeys. In fact the request/response message semantics was designed to satisfy
the following two invariants. We believe that most cache-coherence protocols
would end up obeying these two invariants. With two invariants we can
prove all the pair-wise properties of Section~\ref{pairwise}.

\begin{inv}
\textit{noCross}: A response from a cache to its child for an address cannot
be in transit while another response from the same child to the parent for the
same address is in transit.
\label{noCross}
\end{inv}

\begin{inv}
\textit{respFIFO}: Responses are received in the order in which they are sent,
for a source-destination pair, for the same address.
\label{respFIFO}
\end{inv}

The MSI-based protocol in Figure~\ref{trans} is such that any change in either
the state of a cache or its directory entry takes place only on sending or
receiving a response. The parent upgrades the corresponding directory entry on
sending a response to a child and the child downgrades its state on
sending a response to the parent.

To prove Invariant \fullref{conservative} at some time $t$, consider the case
when a child cache $c$ has sent to or received from the parent $p$ some message
before $t$. Consider the last such message $m$ sent by or received at $c$ at
$t'$. If it is sent rather than received by $c$ at $t'$, then $p$ cannot have
received any message from $c$ after sending $m$ and before $t$ in order to not
violate Invariant \fullref{respFIFO}.  Instead, if $c$ receives $m$ at $t'$,
$p$ will not receive any message after sending $m$ and before $t$ in order to
not violate Invariant \fullref{noCross}. Thus, $p$ will not receive any message
from $c$ after sending or receiving $m$ and before $t$.  Let $t''$ be the time
at which $p$ has sent or received $m$. Whether $p$ sends or receives $m$,
$p.\dstate[c][a]$ at $t''+1$ will be equal to $c.\state[a]$ at $t'+1$ because
the response messages sync up the next-state according to the transitions in
Figure \ref{trans}. Since $p$ cannot receive any message between $t''+1$ and
$t$, it can only upgrade $p.\dstate[c][a]$. $c.\state[a]$ does not change
between $t'+1$ and $t$ as it never changes state. Therefore, Invariant
\fullref{conservative} holds at $t$.

If no message has been sent or received by $c$ before $t$, then $p$ would not
have received any message sent by $c$ before $t$. So $p.\dstate[c][a]$ at time
$t$ is greater than $p.\dstate[c][a]$ at time $0$. Since the initial values of the directory
and the child's state are both set to \In{}, Invariant \textit{conservative}
holds.

To prove Invariant \fullref{localCompatibility}, it is enough to show that a
$c.\state[a]$ is upgraded on receiving a response for $a$ from $c$'s parent $p$
and $p.\dstate[c][a]$ is downgraded on receiving a response for $a$ from $c$.
If this is true, then Invariant \fullref{localCompatibility} will not be
violated on receiving a response. The transitions of Figure \ref{trans} ensures
that Invariant \fullref{localCompatibility} is not violated when sending a
response. Since whenever the state or directory changes, the appropriate
response is sent or received, Invariant \fullref{localCompatibility} holds.

If the value of the state of a child at the time of sending a response is the
same as the value of the directory for that child at the parent at the time of
receiving the response, it follows straightaway that the parent downgrades its
directory on receiving a response from the child, since a child downgrades on
sending a response and the next-states are synced on receiving the response.
The proof for this follows exactly the same arguments presented above to prove
Invariant \fullref{conservative} -- we prove it by considering the last message
sent or received at the child before $t$ (and the case when no message has been
sent or received).

Invariant \fullref{noTransitWrite} can be proved as follows. Consider a response
sent from a parent $p$ at time $t$ received by its child $c$ at time $t'$. No
other response can be received by the parent at time $t_p$ such that $t+1 \le
t_p \le t'$ to not violate Invariant \fullref{noCross}, so $p.\dstate[c][a]$ at
$t+1 \le p.\dstate[c][a]$ at $t'$, since the parent can only upgrade its
directory for $c$. Similarly, no response can be sent by the child at time
$t_c$ such that $t \le t_c \le t'$ for the same reason and hence $c.\state[a]$
at $t \le c.\state[a]$ at $t'$ since the child can only downgrade its state.
This means that during $t_p$ such that $t+1 \le t_p \le t'$, $p.\dstate[c][a]
> \In$ since the parent upgraded its directory for $c$ at $t$. Similarly,
during $t_c$ such that $t \le t_c \le t'$, $c.\state[a] < \Mo$ since the
child upgrades its state at $t'$. Using arguments similar to the proof of
Invariant \fullref{latestValue}, we can establish that no cache can be in \Mo{}
state for address $a$ when a response is in transit from $p$ to $c$.
Similarly, we can prove that no cache can be in \Mo{} state for address $a$
when a response is in transit for address $a$ from $c$ to $p$. Since data is in
transit at the same time as when the response is in transit, Invariant
\fullref{noTransitWrite} follows.

%from the parent and child respectively, it is sufficient to show that the previous state
%or the 
%
%Invariant~\fullref{conservative} can be proved by
%further breaking them into the following invariants between a parent-child pair
%of caches.
%
%\begin{inv}
%\textit{stateChange}:
%\begin{enumerate}
%\item A child sends a response to its parent iff it downgrades its state.
%\item A parent sends a response to a child iff it upgrades the directory of
%that child.
%\item The state and the directory can change only on sending and receiving a
%response; the receiving entity changes its value to that of the sending entity
%immediately after sending the response message.
%\end{enumerate}
%\label{stateChange}
%\end{inv}
%
%\begin{inv}
%\textit{respFIFO}: Responses are received in the order in which they are sent,
%for a source-destination pair, for the same address.
%\label{respFIFO}
%\end{inv}
%
%\begin{inv}
%\textit{noCross}: A response from a cache to its child for an address can not
%be in transit while another response from the same child to the parent for the
%same address is in transit.
%\label{noCross}
%\end{inv}
%
%Invariant~\fullref{conservative} at any time $t$ for child cache $c$ and parent
%$p$ for an address $a$ can be proved as follows. Consider the case when
%$c.\state[a]$ never changed before $t$, then $c.\state[a] = \In$ because of the
%initialization conditions described in Section \ref{System}. No response would
%have been sent by $c$ to its parent for address $a$ before $t$, otherwise
%$c.\state[a]$ would have changed. Thus the parent would not have received any
%response from $c$, $p.\dstate[c][a]$ could have changed only by sending a
%response to $c$ which can only upgrade it from the initial value of $\In$. Thus
%at $t$, $p.\dstate[c][a] \ge c.state[a]$.
%
%Now consider the case when $c.\state[a]$ last changed at $t'$ such that $t' <
%t$ because of receiving a response from $p$, sent at $t''$. $c.\state[a]$ at
%$t'+1 = p.\dstate[c][a]$ at $t''+1$ because of Invariant \fullref{stateChange}.
%$c$ neiter sends nor receives a response at any $t_i$ such that $t' < t_i < t$.
%If $p$ received a response at some $t_m$ such that $t'' < t_m < t$,
%then it would either violate Invariant \fullref{noCross}, or the fact that no
%message was sent between $t' < t_i < t$ from $c$. Thus, $p.\dstate[c][a]$ at $t
%\ge p.\dstate[c][a]$ at $t''+1$ because any message sent can only upgrade
%$p.\dstate[a]$. This ensures that Invariant \fullref{conservative} is obeyed. We
%can similarly prove this invariant holds when $c.\state[a]$ last changes before
%$t$ by sending a response.
%
%Invariant \fullref{stateChange} can be proved directly from the atomic
%transitions for MSI. Other protocols have similar atomic transitions obeying
%these invariants.
%
%Invariant \fullref{respFIFO} can be enforced by the interconnect network which
%ensures a FIFO ordering for messages sent between a pair of nodes.
%
%---------------------------------------------
%
%%Invariant \fullref{respFIFO} is known as point-to-point FIFO ordering in the
%%Network-on-Chip NoC community, and is enforced by common routing algorithms
%%like Dimension-ordered routing \cite{dimension}. Only the parent-to-child
%%channels need to be FIFO ordered, the FIFO ordering in child-to-parent channels
%%can be enforced using the protocol itself.
%
%\begin{inv}
%\textit{noTransitEqual}: When no response is in transit for an address, then
%the state of a cache for that address is the same as its parent's directory for
%that cache and address.
%\label{noTransitEqual}
%\end{inv}
%
%Proving Invariants \fullref{inTransitCondition} and \textit{noTransitEqual},
%needs two further invariants, along with Invariant \fullref{stateChange}.
%
%Invariant \fullref{inTransitCondition} can be proved as follows. Whenever a
%response from a child to the parent is in transit, the parent can not send any
%response to the child, or it will violate Invariant \fullref{noCross}. Using the fact that 
%
%
%
%
%
%
%
%\begin{inv}
%\textit{inTransitCondition}:
%\begin{enumerate}
%\item When a response is sent from a child $c$ at time $t$ for an address $a$,
%and has not been received by the parent $p$ before time $t'$, $\forall t_i, t <
%t_i \le t' @-> c.\state[a]$ at $t_i \le c.\state[a]$ at $t+1$ and $\forall t_i,
%t \le t_i \le t' @->$ $p.\dstate[c][a]$ at $t_i \ge c.\state[a]$ at $t$.
%\item When a response is sent from parent $p$ at time $t$ for an address $a$,
%and has not been received by its child $c$ before time $t'$, $\forall t_i, t <
%t_i \le t' @-> p.\dstate[c][a]$ at $t_i \ge p.\dstate[c][a]$ at $t+1$ and
%$\forall t_i, t \le t_i \le t' @->$ $c.\state[a]$ at $t_i \le p.\dstate[c][a]$ at
%$t$.
%\end{enumerate}
%\label{inTransitCondition}
%\end{inv}
%
%\begin{inv}
%\textit{prevState}: When a response is received from a child to the parent
%(parent to child) for an address, the state of the child for that address at
%the time of sending (receiving) the response is the same as the parent's
%directory for that child and address at the time of receiving (sending) the
%response.
%\label{prevState}
%\end{inv}
%
%
%
%%Top
%%a) conservative: forall p c a t, parent c p -> dir p c a t >= state c a t
%%b) forall p c a t, parent c p -> state p a t >= dir p c a t
%%c) forall c1 c2 a t, sibling c1 c2 -> compatible c1 c2 a t
%%d) nextClean: ~ clean c a t -> clean c a (S t) -> exists s d ts, recv data s a t d /\ send data s a ts d /\ clean s a ts /\ ts <= t /\ forall ti, ts <= ti < t ->  ~ write a ti
%%
%%(from=state(t1)=dir(t2) -> betweenPSendCRecv:dir(t)>=to&state(t)<=from, betweenCSendPRecv:state(t)<=to&dir(t)>=from), noMesgImp:state(t)=dir(t) -> conservative (ignore from=, use only a, b)
%%a) Child sends message iff it downgrades
%%b) Parent sends message iff it upgrades
%%c) State is changed only on sending or receiving a message
%%d) FIFO messages
%%e) cross: Message from child and parent do not cross
%%
%%cross:
%%a) parent send message only on receiving request
%%b) req from child does not overtake response from child (prove using the from states)
%%c) resp from parent does not overtake request from parent
%%d) voluntary response sent from child only when not waiting
%%
%%In this section we specify a set of important invariants for a parent-child pair of caches.

%%  LocalWords:  Invariants invariants