### 2014-02-06 paper/Abstract.tex
this the the correct version

%Correct implementation of cache coherent systems is of high importance in
%industrial use.
We provide a precise formal definition of a cache coherent system as one
which obeys store atomicity. We give a Coq-based proof of a directory-based MSI cache
coherence protocol for an infinite family of cache hierarchies. We specify a
set of invariants which imply store atomicity, enabling other protocols to be
verified for store atomicity automatically as long as they satisfy the
invariants. Our key insight for proving these invariants is a set of
fundamental properties which are shared with other protocols like MESI, MOSI
and MOESI, enabling reuse of portions of our MSI proof for their verification.

%Previously, it has been shown
%that these atomic transitions can be translated automatically into efficient
%hardware. Verifying the specification based on atomic transitions therefore
%guarantees the correctness of the actual implementation.
%The proof is built in a modular fashion, identifying key invariants common with
%other protocols like MESI, MOSI, \etc. We informally show how these other
%protocols can be proved reusing modules from the proof of the MSI protocol.