### 2014-02-06 paper/Title.tex

%\begin{titlepage}

\title{Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies}

% hierarchical cache coherence
% modular proof

%       {\Large (Regular Submission)}}

%\author{Names withheld for submission}%\vspace{30pt}}

%% \author{
%% %  \adamc{Fill in authors, including Richard or any other appropriate additions}
%%    Muralidaran Vijayaraghavan\\Massachusetts Institute of Technology\\vmurali@csail.mit.edu \and
%%    Nirav Dave\\SRI International\\ndave@csl.sri.com \and
%%    Adam Chlipala\\Massachusetts Institute of Technology\\adamc@csail.mit.edu \and
%%    Arvind\\Massachusetts Institute of Technology\\arvind@csail.mit.edu
%% }

\author{Muralidaran Vijayaraghavan\inst{1} \and Nirav Dave\inst{2} \and Adam Chlipala\inst{1} \and Arvind\inst{1}}

\institute{{Massachusetts Institute of Technology,}\\{Computer Science and
    Artificial Intelligence Laboratory, Cambridge, MA, USA}\\\email{\{vmurali, adamc, arvind\}@csail.mit.edu} \and
           {SRI International, Computer Science Laboratory, Menlo Park, CA,
       USA}\\\email{ndave@csl.sri.com}}

\date{}

\maketitle % \thispagestyle{empty}

%=========================================================================
%  Abstract
%=========================================================================

\begin{abstract}
\input{Abstract}
\end{abstract}

%\bigskip

\centerline{\footnotesize {\bf Keywords}: Cache coherence, proof assistant, cache hierarchy, guarded atomic actions}

%\end{titlepage}