### 2014-02-06 paper/RelatedWork.tex
stuff

\section{Related work}
\label{relatedWork}
These are lifted as is from that guy. Need to figure out how to restate them, and make it more concise.

For verification of high level specifications, modern industrial practice
consists of modeling small instances of the protocols, e.g., three CPUs
handling two addresses and one bit of data, in terms of interleaving atomic
steps, in guard/action languages such as Murphi \cite{tla} or TLA+ \cite{tla}, and
exploring the reachable states through explicit state enumeration.
Symbolic methods, e.g., BDD [31] or SAT [129, 142], are yet to succeed for the
verification of cache coherence protocols.
Monolithic formal verification methods – methods that treat the protocol as a whole – have
been used fairly routinely for verifying cache coherence protocols from the early 1990s
[10 (2 nodes, murphi), 61 (), 141]. However, these monolithic techniques will not be able to handle the very
large state space of hierarchical protocols. Compositional verification techniques are essential.