### 2014-02-06 paper/RelatedWork.tex

\section{Related work}
\label{relatedWork}

Formal verification of hardware-oriented cache coherence protocols, is mainly
done using model checking. In practice, explicit state model checking tools like
Murphi \cite{murphi} or TLC \cite{tlc} are only able to handle small and
simplified systems, \eg{} a single-level cache hierarchy with three CPUs, two
addresses, and one bit of data. Symbolic model checking techniques have fared
slightly better; McMillan \etal{} have verified a 2-level MSI protocol based on
Gigamax distributed multiprocessor using SMV.  Optimizations on these techniques
(\eg{} partial order reduction \cite{part}, symmetry reduction \cite{sym1,
  sym2}, compositional reasoning \cite{somethink}, \etc{} scale the approach
further, but they are still unable to handle realistic hierarchical protocols.

Theorem provers have also been used in verifying cache coherence protocols. The
FLASH coherence protocol \cite{flash} was verified using Aggregation of
Distributed transactions \cite{park}, where several state transitions accessing
localized state intuitively correspond to an aggregate atomic transition
accessing all global state. A proof of a mapping between the local transitions
and the agreggate one must be provided. Once this is done, a model checker can
be used to verify the aggregate model is correct. However, We believe that the
verification challenge here is precisely the proof of such a mapping which was
not discussed.

%These are lifted as is from that guy. Need to figure out how to restate them, and make it more concise.
%
%For verification of high level specifications, modern industrial practice
%consists of modeling small instances of the protocols, e.g., three CPUs
%handling two addresses and one bit of data, in terms of interleaving atomic
%steps, in guard/action languages such as Murphi \cite{tla} or TLA+ \cite{tla}, and
%exploring the reachable states through explicit state enumeration.
%Symbolic methods, e.g., BDD [31] or SAT [129, 142], are yet to succeed for the
%verification of cache coherence protocols.
%Monolithic formal verification methods – methods that treat the protocol as a whole – have
%been used fairly routinely for verifying cache coherence protocols from the early 1990s
%[10 (2 nodes, murphi)]. However, these monolithic techniques will not be able to handle the very
%large state space of hierarchical protocols. Compositional verification techniques are essential. 
%
%For coherence protocols which only consider one level (or nonhierarchical protocols),
%various techniques have been proposed for the verification. These techniques basically
%include explicit state enumeration \cite{sci-cache-coherence} and symbolic model checking \cite{gigamax}
%Opti-
%mizations on these techniques include partial order reduction [24], symmetry reduction
%[25, 72], compositional reasoning [96], predicate abstraction [22, 49, 82], etc. There is
%also a large body of work on parameterized verification [42, 53, 54, 100, 117]. However,
%none of them have shown the capability to verify hierarchical coherence protocols with
%realistic complexity.
%
%McMillan et al. [97, 102] modeled a 2-level MSI coherence protocol based on the
%Gigamax distributed multiprocessor [102]. In the protocol, bus snooping is used in both
%levels. SMV [97] was used to verify this protocol with two clusters each having six
%processors, for both safety and liveness properties. In terms of the complexity, many
%realistic details are abstracted away from the protocol so that the protocol does not have
%the typical complexity that a hierarchical coherence protocol has.
%