### 2014-02-06 paper/RelatedWork.tex
stuff

\section{Related Work}
\label{relatedWork}

Formal verification of hardware-oriented cache coherence protocols is mainly
done using model checking. In practice, explicit-state model checking tools like
Murphi \cite{murphi} or TLC \cite{tlc} are only able to handle small and
simplified systems, \eg{} a single-level cache hierarchy with three CPUs, two
addresses, and one bit of data. Symbolic model checking techniques have fared
slightly better; McMillan \etal{} have verified a 2-level MSI protocol based on
the Gigamax distributed multiprocessor using SMV \cite{gigamax}.  Optimizations on these techniques
(\eg{} partial order reduction \cite{part}, symmetry reduction \cite{sym1,
  sym2}, compositional reasoning \cite{comp}, \etc{}) scale the approach
further, but they are still unable to handle realistic hierarchical protocols.

Theorem provers have also been used in verifying cache coherence protocols. The
FLASH coherence protocol was verified using aggregation of
distributed \adamc{why are these words capitalized?} transactions \cite{park},
where several state transitions accessing localized state intuitively
correspond to an aggregate atomic transition accessing all global state. A
proof of a mapping between the local transitions and the agreggate one must be
provided, at which point a model checker can be used to verify that the
aggregate model is correct. However, we believe that the verification challenge
here is precisely the proof of such a mapping, which was not discussed.

\xxx{MUST WRITE ABOUT XIAOWEI's CACHET and JOE STOY'S PVS proof}

%These are lifted as is from that guy. Need to figure out how to restate them, and make it more concise.
%
%For verification of high level specifications, modern industrial practice
%consists of modeling small instances of the protocols, e.g., three CPUs
%handling two addresses and one bit of data, in terms of interleaving atomic
%steps, in guard/action languages such as Murphi [50] or TLA+ [84], and
%exploring the reachable states through explicit state enumeration.
%Symbolic methods, e.g., BDD [31] or SAT [129, 142], are yet to succeed for the
%verification of cache coherence protocols.
%Monolithic formal verification methods – methods that treat the protocol as a whole – have
%been used fairly routinely for verifying cache coherence protocols from the early 1990s
%[10 (2 nodes, murphi)]. However, these monolithic techniques will not be able to handle the very
%large state space of hierarchical protocols. Compositional verification techniques are essential. 
%
%For coherence protocols which only consider one level (or nonhierarchical protocols),
%various techniques have been proposed for the verification. These techniques basically
%include explicit state enumeration [133] and symbolic model checking [102]
%Opti-
%mizations on these techniques include partial order reduction [24], symmetry reduction
%[25, 72], compositional reasoning [96], predicate abstraction [22, 49, 82], etc. There is
%also a large body of work on parameterized verification [42, 53, 54, 100, 117]. However,
%none of them have shown the capability to verify hierarchical coherence protocols with
%realistic complexity.
%
%McMillan et al. [97, 102] modeled a 2-level MSI coherence protocol based on the
%Gigamax distributed multiprocessor [102]. In the protocol, bus snooping is used in both
%levels. SMV [97] was used to verify this protocol with two clusters each having six
%processors, for both safety and liveness properties. In terms of the complexity, many
%realistic details are abstracted away from the protocol so that the protocol does not have
%the typical complexity that a hierarchical coherence protocol has.
%